---
title: "BT_mating"
author: "Jeremias Brand"
date: "4 August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, results="hide", warning=FALSE}
source("R/functions.R")
# pacman::p_load(ape, phytools, phangorn, purrr, tidyr, dplyr, ggplot2, ggtree, hisse, RColorBrewer, readxl, lattice, writexl, grid, gridExtra, gtable, ggstance))
pkg <- c("ape", "phytools", "phangorn", "purrr", "tidyr", "dplyr", "ggplot2", "ggtree", "hisse", "RColorBrewer", "readxl", "lattice", "writexl", "grid", "gridExtra", "gtable", "ggstance", "coda", "foreach", "cowplot", "data.table")
lapply(pkg, require, character.only = TRUE)

```


```{r}
stats_all <- fread("output/06_BT_statistics_all_runs.csv", stringsAsFactors = FALSE)
quant_all <- read.csv("output/06_BT_quantiles_all_runs.csv", stringsAsFactors = FALSE)
Lh_all <- read.csv("output/06_BT_BF_all_runs.csv", stringsAsFactors = FALSE)
```

## We want to vizualize the transition rates for all models

There are 27 separate tests so we should ideally not have to plot them completely manually


```{r}
v <- unique(stats_all$varnames)
t <- unique(stats_all$tree)
p <- unique(stats_all$prior)
```


```{r}
library(diagram)
varN <- unique(stats_all$varnames)
display_names <- c("sperm: antrum only/some hypodermic 1/0\n long bristle/reduced bristle 1/0",
                   "sperm: antrum only/hypodermic sperm present 1/0\n antrum thick/antrum simple 1/0",
                   "long bristle/reduced bristle 1/0\nantrum thick/antrum simple 1/0")
trees <- unique(stats_all$tree)
tree_names <- c("H-IQ-TREE", "H-ExaBayes", "C-IQ-TREE")
priors <- unique(stats_all$prior)
stats_all$Mean <- round(stats_all$Mean,1)
stats_all <- data.table(stats_all)

transitions <- c("q12","q13","q21","q24","q31","q34","q42","q43")
# three trees and three priors so we should have a 3x3 grid
pdf("fig/06_Transition_matrix_raw_v02.pdf", height = 10, width = 7, paper = "a4")
par(mfrow=c(3,3),
    mar=c(0,0,5,0))
for (i in 1:length(varN)){
  for (j in 1:length(trees)){
    for (k in 1:length(priors)){
    model <- stats_all[varnames == varN[[i]] &
                         tree == trees[j] &
                         prior == priors[k] &
                         model == "dependent" &
                       .rownames %in% transitions]
    
    BF <- data.table(Lh_all)[VARNAMES == varN[i] &
                         t_name == trees[j] &
                         prN == priors[k], "marginalLH_BF"]
    
    LH_ind <- data.table(Lh_all)[VARNAMES == varN[i] &
                         t_name == trees[j] &
                         prN == priors[k], "mean_marginalLH_independent"]
    
    LH_dep <- data.table(Lh_all)[VARNAMES == varN[i] &
                         t_name == trees[j] &
                         prN == priors[k], "mean_marginalLH_dependent"]
                 
    trs <- model[,"Mean"]
    trs_num <- scales::rescale(as.numeric(trs$Mean), to=c(1,5))
    nms <- c("00","01","10","11")
    m <- matrix(
    c(0,trs[1],trs[2],0,
    trs[3],0,0,trs[4],
    trs[5],0,0,trs[6],
    0,trs[7],trs[8],0),
    nrow = 4, dimnames = list(nms,nms))
    # make an arrow sclaing vector
    m_arr <- matrix(
    c(0,trs_num[1],trs_num[2],0,
    trs_num[3],0,0,trs_num[4],
    trs_num[5],0,0,trs_num[6],
    0,trs_num[7],trs_num[8],0),
    nrow = 4)
    mn <- apply(m,1,as.numeric)
    apply(mn,MARGIN=1, scales::rescale)
        Lines <- list(bquote(.(tree_names[j])),
                  bquote(paste("Lh" ["ind"], ":", .(LH_ind[[1]]))),
                  bquote(paste("Lh" ["dep"], ":", .(LH_dep[[1]]))),
                  bquote(paste("BF:", .(BF[[1]]))))  
    plotmat(m, pos = c(2,2), curve = 0.05,
        box.type = "rect", box.prop = 0.5, 
        shadow.size = 0,
        arr.lwd = 1, arr.width = (m_arr * 0.04)+0.05,
        arr.length = m_arr * 0.1, arr.type = "triangle",
         main = mtext(do.call(expression, Lines),side=3, line=c(2,1,0,-1),
                     cex = 0.7))
    
    }
  }
}
dev.off()
# display_names[i],"\n",



pdf("fig/06_Transition_matrix_scaled_v02.pdf", height = 10, width=9)
par(mfrow=c(3,3),
    mar=c(0,0,5,0))
for (i in 1:length(varN)){
  for (j in 1:length(trees)){
    for (k in 1:length(priors)){
    model <- stats_all[varnames == varN[i] &
                         tree == trees[j] &
                         prior == priors[k] &
                         model == "dependent" &
                       .rownames %in% transitions]
    BF <- round(data.table(Lh_all)[VARNAMES == varN[i] &
                         t_name == trees[j] &
                         prN == priors[k], "marginalLH_BF"], 1)[[1]]
    
    LH_ind <- round(data.table(Lh_all)[VARNAMES == varN[i] &
                         t_name == trees[j] &
                         prN == priors[k], "mean_marginalLH_independent"],1)[[1]]
    
    LH_dep <- round(data.table(Lh_all)[VARNAMES == varN[i] &
                         t_name == trees[j] &
                         prN == priors[k], "mean_marginalLH_dependent"],1)[[1]]
                 
    trs <- model[,"Mean"]
    trs_num <- scales::rescale(as.numeric(trs$Mean), to=c(1,5))
    trs <- round(scales::rescale(as.numeric(trs$Mean), to = c(0.01, 1)),2)
    nms <- c("00","01","10","11")
    m <- matrix(
    c(0,trs[1],trs[2],0,
    trs[3],0,0,trs[4],
    trs[5],0,0,trs[6],
    0,trs[7],trs[8],0),
    nrow = 4, dimnames = list(nms,nms))
    # make an arrow sclaing vector
    m_arr <- matrix(
    c(0,trs_num[1],trs_num[2],0,
    trs_num[3],0,0,trs_num[4],
    trs_num[5],0,0,trs_num[6],
    0,trs_num[7],trs_num[8],0),
    nrow = 4)
    mn <- apply(m,1,as.numeric)
    apply(mn,MARGIN=1, scales::rescale)
    Lines <- list(bquote(.(tree_names[j])),
                  bquote(paste("Lh" ["ind"], ":", .(LH_ind[[1]]))),
                  bquote(paste("Lh" ["dep"], ":", .(LH_dep[[1]]))),
                  bquote(paste("BF:", .(BF[[1]]))))         
    
    par(mar = c(2, 2, 2, 2))

plotmat(m, pos = c(2,2), curve = 0.05,
        box.type = "rect", box.prop = 0.5, 
        shadow.size = 0,
        arr.lwd = 1, arr.width = (m_arr * 0.04)+0.05,
        arr.length = m_arr * 0.1, arr.type = "triangle",
        main = mtext(do.call(expression, Lines),side=3, line=c(1,0,-1,-2),
                     cex = 0.7))
    }
  }
}
dev.off()
```

#######################
# Plotting the posterior distributions


```{r}
options(scipen=99999)
BTV <- "BayesTraitsV3"
BT_prefix <- "BT_19-07"
OUTDIR = "BT/BT_19-07"
stones_no = 1000
stones_iter = 10000
burnin = 10000000
iterations = 510000000
sampelfreq = 100
chains = 4
priors = c("PriorAll uniform 0 100", "PriorAll exp 10", "RJHP gamma 0 1 0 1")
options(scipen=99999)
varpairs <- list(c("Rec_binary", "bristle_binary"),
                 c("Rec_binary", "an_binary"),
                 c("bristle_binary", "an_binary"))
```
## Reading results in again.

from the log files we want to extract thr posterior samples
from the stones file we want just the marginal likelyhood (the likelyhood of the model given the data)

```{r}
parse_BT_output <- function(trees, priors, OUTDIR, varpairs) {
  # given files in specific format this parses them and returns the result
  var1 = varpairs[1]
  var2 = varpairs[2]
  tre_ls <- list()
  print(paste0(c("Starting parsing of traits:", var1, var2)))
for (i in 1:length(trees)){
  t = trees[[i]]
  t_name <- names(trees)[[i]]
  print(paste("Working on tree:",t_name))
  pr_ls <- list()
  for (p in 1:length(priors)) {
      prN <- unlist(strsplit(priors[p], split = " "))[2]
      print(prN)
      s_in = c(); c_in = mcmc.list(); c_dep = mcmc.list(); s_dep = c()
      for (c in 1:chains){
      n_in <- paste0(OUTDIR,"/",BT_prefix,"_",var1,"_",var2, "_", t_name,"_", "indep_",prN,"_ch", c)
      logfile_in <- paste0(n_in, ".Log.txt")
      stonefile_in <- paste0(n_in, ".Stones.txt")
      n_dep <- paste0(OUTDIR,"/",BT_prefix,"_",var1,"_",var2, "_", t_name,"_", "dep_",prN,"_ch", c)
      logfile_dep <- paste0(n_dep, ".Log.txt")
      stonefile_dep <- paste0(n_dep, ".Stones.txt")
      c_in[[c]] <- read_bt_discrete_mcmc(logfile_in)
      c_dep[[c]] <- read_bt_discrete_mcmc(logfile_dep)
      # stones results
      s_in <- c(s_in,read_bt_discrete_marginalLH(stonefile_in))
      s_dep <- c(s_dep,read_bt_discrete_marginalLH(stonefile_dep))
      }
      pr_res <- setNames(list(c_in,c_dep,s_in,s_dep),
                         paste(rep(prN), c("chain_in", "chain_dep", "stone_in", "stone_dep")))
      pr_ls[[p]] <- pr_res
      names(pr_ls)[[p]] <- prN
  }
  tre_ls[[i]] <- pr_ls
  names(tre_ls)[[i]] <- t_name
  }
  return(tre_ls)
}

read_bt_discrete_marginalLH <- function(path) {
  require(coda)
  require(dplyr)
  BT_file <- path
  FullOut <-  scan(file = BT_file, what="c", quiet=T, sep="\n")
  BT_log <- read.table(BT_file, skip = (grep("Log marginal likelihood:", FullOut) - 1), sep = "\t", header = FALSE, quote="\"") %>% data.frame()
  return(BT_log$V2)
}

read_bt_discrete_mcmc <- function(path) {
  require(coda)
  require(dplyr)
  library(data.table)
  BT_file <- path
  FullOut <-  scan(file = BT_file, what="c", quiet=T, sep="\n")
  BT_log <- fread(BT_file, skip = (grep("Iteration	Lh", FullOut) - 1), sep = "\t", header = TRUE, quote="\"") %>% data.frame()
  BT_log <- BT_log[,1:ncol(BT_log)-1]
  BT_log <- select(BT_log, -Tree.No, -Iteration)
  BT_log_mcmc <- mcmc(BT_log)
  return(BT_log_mcmc)
}

read_bt_discrete_mcmc2 <- function(path) {
  require(coda)
  require(dplyr)
  library(data.table)
  BT_file <- path
  FullOut <-  scan(file = BT_file, what="c", quiet=T, sep="\n")
  BT_log <- fread(BT_file, skip = (grep("Iteration	Lh", FullOut) - 1), sep = "\t", header = TRUE, quote="\"") %>% data.frame()
  BT_log <- BT_log[,1:ncol(BT_log)-1]
  BT_log <- tryCatch(
    {select(BT_log, -Model.string, -`Dep...InDep`, -Tree.No, -Iteration)},
                     error=function(cond){
                       return(select(BT_log, -Model.string, -Tree.No, -Iteration))
                     })
  BT_log_mcmc <- mcmc(BT_log)
  return(BT_log_mcmc)
}

parse_BT_output2 <- function(trees, priors, OUTDIR, varpairs) {
  # given files in specific format this parses them and returns the result
  var1 = varpairs[1]
  var2 = varpairs[2]
  tre_ls <- list()
  print(paste0(c("Starting parsing of traits:", var1, var2)))
for (i in 1:length(trees)){
  t = trees[[i]]
  t_name <- names(trees)[[i]]
  print(paste("Working on tree:",t_name))
  pr_ls <- list()
  for (p in 1:length(priors)) {
      prN <- priors[[p]]
      print(prN)
      s_in = c(); c_in = mcmc.list(); c_dep = mcmc.list(); s_dep = c()
      for (c in 1:chains){
      n_in <- paste0(OUTDIR,"/",BT_prefix,"_",var1,"_",var2, "_", t_name,"_", "indep_",prN,"_ch", c)
      logfile_in <- paste0(n_in, ".Log.txt")
      stonefile_in <- paste0(n_in, ".Stones.txt")
      n_dep <- paste0(OUTDIR,"/",BT_prefix,"_",var1,"_",var2, "_", t_name,"_", "dep_",prN,"_ch", c)
      logfile_dep <- paste0(n_dep, ".Log.txt")
      stonefile_dep <- paste0(n_dep, ".Stones.txt")
      c_in[[c]] <- read_bt_discrete_mcmc2(logfile_in)
      c_dep[[c]] <- read_bt_discrete_mcmc2(logfile_dep)
      # stones results
      print(deparse(substitute(stonefile_in)))
      s_in <- c(s_in,read_bt_discrete_marginalLH(stonefile_in))
      s_dep <- c(s_dep,read_bt_discrete_marginalLH(stonefile_dep))
      }
      pr_res <- setNames(list(c_in,c_dep,s_in,s_dep),
                         paste(rep(prN), c("chain_in", "chain_dep", "stone_in", "stone_dep")))
      pr_ls[[p]] <- pr_res
      names(pr_ls)[[p]] <- prN
  }
  tre_ls[[i]] <- pr_ls
  names(tre_ls)[[i]] <- t_name
  }
  return(tre_ls)
}
```

```{r}

all_merged <- readRDS(file = paste0(OUTDIR,"/all_merged.Rdata"))
str(all_merged[[1]])
names(all_merged) <- sapply(varpairs, paste0, collapse="_v_")
```


This produces a nested list:

* tree/data used
  + prior used
    - independent chains (4 mcmc lists)
    - dependent chains (4 mcmc lists)
    - marginal likelyhoods independent (vector)
    - marginal likelihoods dependent (vector)

We want to:
plot posterior distribution depending on prior
conduct likelihood ratio tests and store this as well.

```{r}
# foreach(x = 1:length(all_parsed)) %dopar% {
#   lapply(pkg, require, character.only = TRUE)
#   VARNAMES <- names(all_parsed)[x]
#   tre_ls <- all_parsed[[x]]
#   sink(paste0(OUTDIR,"/diagnostics/gelman_diagnostics",VARNAMES ,".txt"))
# for (i in 1:ltre){
#   t_name <- names(trees)[[i]]
#   for (p in 1:length(priors)) {
#     print(t_name)
#     print(priors[[p]])
#     print(paste(VARNAMES, "Gelman diagnostics for independent run"))
#     print(gelman.diag(tre_ls[[i]][[p]][[1]]))
#     print(paste(VARNAMES, "Gelman diagnostics for dependent run"))
#     print(gelman.diag(tre_ls[[i]][[p]][[2]]))
#     print("=====================================================================")
#     }
# }
# sink()
# 
# 
# pdf(paste0(OUTDIR,"/diagnostics/gelman_diagnostics",VARNAMES ,".pdf"), width = 21, height = 14)
#   par(mfrow=c(4,6),oma = c(0, 0, 2, 0))
#   for (i in 1:ltre){
#     for (p in 1:length(priors)) {
#       gelman.plot(tre_ls[[i]][[p]][[1]], auto.layout = FALSE, autoburnin=FALSE)
#       plot.new()
#       title(paste(VARNAMES, names(trees)[[i]], priors[[p]], "Gelman diagnostics for independent (1st) and dependent (2nd) run"), outer = TRUE, cex = 1.5)
#       plot.new()
#       gelman.plot(tre_ls[[i]][[p]][[2]], auto.layout = FALSE, autoburnin=FALSE)
#       }
#   }
# dev.off()
#   
# }

```


## Same for rjhp

```{r}
priors = c("uniform 0 100", "exp 10", "rjhp")

for (x in  1:length(all_merged)) {
  #for (x in 1:length(all_merged)) {
  lapply(pkg, require, character.only = TRUE)
  VARNAMES <- names(all_merged)[x]
  tre_ls <- all_merged[[x]]
  sink(paste0(OUTDIR,"/diagnostics/gelman_diagnostics_",VARNAMES ,".txt"))
for (i in 1:ltre){
  t_name <- names(trees)[[i]]
  for (p in 1:length(priors)) {
    print(t_name)
    print(priors[[p]])
    print(paste(VARNAMES, "Gelman diagnostics for independent run"))
    print(gelman.diag(tre_ls[[i]][[p]][[1]]))
    print(paste(VARNAMES, "Gelman diagnostics for dependent run"))
    print(gelman.diag(tre_ls[[i]][[p]][[2]]))
    print("=====================================================================")
    }
}
sink()


pdf(paste0(OUTDIR,"/diagnostics/gelman_diagnostics_",VARNAMES ,".pdf"), width = 21, height = 14)
  par(mfrow=c(4,6),oma = c(0, 0, 2, 0))
  for (i in 1:ltre){
    for (p in 1:length(priors)) {
      gelman.plot(tre_ls[[i]][[p]][[1]], auto.layout = FALSE, autoburnin=FALSE)
      plot.new()
      title(paste(VARNAMES, names(trees)[[i]], priors[[p]], "Gelman diagnostics for independent (1st) and dependent (2nd) run"), outer = TRUE, cex = 1.5)
      plot.new()
      gelman.plot(tre_ls[[i]][[p]][[2]], auto.layout = FALSE, autoburnin=FALSE)
      }
  }
dev.off()
  
}
```


After the diagnostic I plot the posterior distributions

```{r}
library(cowplot)
ltre = 3
priors = c("uniform 0 100", "exp 10", "rjhp")
prtitle = c("Uniform prior", "Exponential prior", "Reversible-jump hyperprior")
titles <- c("Received sperm location | Bristle state,", "Received sperm location | Antrum state,", "Bristle state | Antrum state,")
treetitle <- c("H-IQ-TREE,", "H-ExaBayes,", "C-IQ-TREE,")

for(x in 1: length(all_merged))  { # BT pairs
  stats_mcmcs <- list()
  quantile_mcmcs <- list()
  VARNAMES <- names(all_merged)[x]
  tre_ls <- all_merged[[x]]
  
  pdf(paste0("fig/06_BT_posterior_",VARNAMES ,"_fancy2.pdf"), width = 14, height = 14)
  for (i in 1:ltre){
    t = trees[[i]]
    t_name <- names(trees)[[i]]
    # keep priors together
    plts <- list()
    for (p in 1:length(priors)) {
      prN <- unlist(strsplit(priors[p], split = " "))[2]
      mcmcs <- tre_ls[[i]][[p]][[2]]
      df_mcmcs <- as.data.frame(do.call(rbind,mcmcs))
      
      # make it small for testing:
      #nrow(df_mcmcs)
      #df_mcmcs <- df_mcmcs[1:10000,]
      
      df_tidy_all <- df_mcmcs %>%
        rename(Log_likelihood = Lh,
               '00 to 01' = q12,
               '00 to 10' = q13,
               '01 to 00' = q21,
               '01 to 11' = q24,
               '10 to 00' = q31,
               '10 to 11' = q34,
               '11 to 01' = q42,
               '11 to 10' = q43
               ) %>%
        select(Log_likelihood, 
               '00 to 10',
               '01 to 11',
               '00 to 01',
               '10 to 11',
               '10 to 00',
               '11 to 01',
               '01 to 00',
               '11 to 10')
      
      df_tidy <- df_tidy_all %>%
        select(-Log_likelihood) %>%
        gather(key = var, value = val) %>%
        mutate(var = factor(var, levels = c(
          "00 to 10","00 to 01","10 to 00","01 to 00",
          "01 to 11","10 to 11","11 to 01","11 to 10"
        )))
      
      pLH <- ggplot(df_tidy_all) +
        geom_histogram(aes(x=Log_likelihood, y=..ncount..), bins = 50) +
        #geom_histogram(aes(x=val, y=(..count../sum(..count..))), bins= 50) +
        #facet_wrap(~var,nrow= 2) +
        #ggtitle(paste(title[x], treetitle[i], prtitle[p])) +
        ylab("Frequency") +
        xlab("Likelihood") +
        theme_bw()
      
      p2 <- ggplot(df_tidy) +
        geom_histogram(aes(x=val, y=..ncount..), bins = 50) +
        #geom_histogram(aes(x=val, y=(..count../sum(..count..))), bins= 50) +
        facet_wrap(~var,nrow= 2) +
        #ggtitle(paste(title[x], treetitle[i], prtitle[p])) +
        ylab("Frequency") +
        xlab("Rate") +
        theme_bw()

      plot_row <- cowplot::plot_grid(p2, pLH)
  # now add the title
  title <- ggdraw() + 
  draw_label(
    paste(titles[x], treetitle[i], prtitle[p]),
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
  plts[[p]] <- plot_grid(
  title, plot_row,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
  )
  
    }
    
  print(plot_grid(plts[[1]], plts[[2]], plts[[3]], ncol=1))
  }
  
  
  
  
  
  dev.off()
}

```


```{r}
priors = c("uniform 0 100", "exp 10", "rjhp")
stats_all <- list()
#stats_all <- foreach(x = 1: length(all_merged)) %dopar% {
  for(x in 1:length(all_merged)){
  stats_mcmcs <- list()
  quantile_mcmcs <- list()
  lapply(pkg, require, character.only = TRUE)
  VARNAMES <- names(all_merged)[x]
  tre_ls <- all_merged[[x]]

  for (i in 1:ltre){
    t = trees[[i]]
    t_name <- names(trees)[[i]]
    for (p in 1:length(priors)) {
      prN <- unlist(strsplit(priors[p], split = " "))[1]
      mcmcs <- tre_ls[[i]][[p]][[1]]
      
      s_mcmcs <- summary(mcmcs)
      s_stats <- broom::tidy(s_mcmcs$statistics)
      s_stats$tree <- t_name
      s_stats$varnames <- VARNAMES
      s_stats$prior <- prN
      s_stats$model <- "independent"
      stats_mcmcs <- rbind(stats_mcmcs, s_stats)

      mcmcs <- tre_ls[[i]][[p]][[2]]
      s_mcmcs <- summary(mcmcs)
      s_stats <- broom::tidy(s_mcmcs$statistics)
      s_stats$tree <- t_name
      s_stats$varnames <- VARNAMES
      s_stats$prior <- prN
      s_stats$model <- "dependent"
      stats_mcmcs <- rbind(stats_mcmcs, s_stats)
      }
  }
  stats_all <- append(stats_all, list(stats_mcmcs))
}

quant_all <- list()
#quant_all <- foreach(x = 1: length(all_parsed)) %dopar% {
  for(x in 1:length(all_merged)){
  stats_mcmcs <- list()
  quantile_mcmcs <- list()
  lapply(pkg, require, character.only = TRUE)
  VARNAMES <- names(all_merged)[x]
  tre_ls <- all_merged[[x]]

  for (i in 1:ltre){
    t = trees[[i]]
    t_name <- names(trees)[[i]]
    for (p in 1:length(priors)) {
      prN <- unlist(strsplit(priors[p], split = " "))[1]
      
      mcmcs <- tre_ls[[i]][[p]][[1]]
      s_mcmcs <- summary(mcmcs)
      s_quant <- broom::tidy(s_mcmcs$quantiles)
      s_quant$tree <- t_name
      s_quant$varnames <- VARNAMES
      s_quant$prior <- prN
      s_quant$model <- "independent"
      quantile_mcmcs <- rbind(quantile_mcmcs, s_quant)

      mcmcs <- tre_ls[[i]][[p]][[2]]
      s_mcmcs <- summary(mcmcs)
      s_quant <- broom::tidy(s_mcmcs$quantiles)
      s_quant$tree <- t_name
      s_quant$varnames <- VARNAMES
      s_quant$prior <- prN
      s_quant$model <- "dependent"
      quantile_mcmcs <- rbind(quantile_mcmcs, s_quant)
      }
  }
  quant_all <- append(quant_all, list(quantile_mcmcs))
}

stats_all <- do.call(rbind, stats_all)
quant_all <- do.call(rbind, quant_all)

write.csv(stats_all, "output/06_BT_statistics_all_runs.csv", row.names = FALSE)
write.csv(quant_all, "output/06_BT_quantiles_all_runs.csv", row.names = FALSE)
```

## Extract Lh and calculate BF

```{r}
priors = c("uniform 0 100", "exp 10", "rjhp")
Lh_all <- list()
#Lh_all <- foreach(x = 1: length(all_merged)) %dopar% {
 for(x in 1:length(all_merged)){
  Lh_df <- data.frame(VARNAMES = NA,
                      t_name= NA, prN = NA,
                     harmonic_mean_LH_independent=NA, harmonic_mean_LH_dependent=NA,
        mean_marginalLH_independent=NA, mean_marginalLH_dependent=NA, marginalLH_BF=NA)
  lapply(pkg, require, character.only = TRUE)
  VARNAMES <- names(all_merged)[x]
  tre_ls <- all_merged[[x]]

  for (i in 1:ltre){
    t = trees[[i]]
    t_name <- names(trees)[[i]]
    for (p in 1:length(priors)) {
      prN <- unlist(strsplit(priors[p], split = " "))[1]
      
      harmonic_mean_LH_independent <- mean(unlist(tre_ls[[i]][[p]][[1]][,"Lh"]))
      harmonic_mean_LH_dependent <-   mean(unlist(tre_ls[[i]][[p]][[2]][,"Lh"]))
      mean_marginalLH_independent <- mean(tre_ls[[i]][[p]][[3]])
      mean_marginalLH_dependent <- mean(tre_ls[[i]][[p]][[4]])
      
      marginalLH_BF = 2 * (mean_marginalLH_dependent - mean_marginalLH_independent)
      
      Lh_df <- rbind(Lh_df, c(VARNAMES, t_name, prN,
                     harmonic_mean_LH_independent, harmonic_mean_LH_dependent,
        mean_marginalLH_independent, mean_marginalLH_dependent, marginalLH_BF))
      }
  }
  Lh_all <- append(Lh_all,list(Lh_df))
}

Lh_all <- na.omit(do.call(rbind,Lh_all))
write.csv(Lh_all, "output/06_BT_BF_all_runs.csv", row.names = FALSE)
```

