---
title: "SA_exploration_LUK"
output: pdf_document
---

```{r setup, include=FALSE}
require(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = "..")
```

```{r message=FALSE, results="hide", warning=FALSE}
source("R/functions.R")
# pacman::p_load(ape, phytools, phangorn, purrr, tidyr, dplyr, ggplot2, ggtree, hisse, RColorBrewer, readxl, lattice, writexl, grid, gridExtra, gtable, ggstance))
pkg <- c("ape", "phytools", "phangorn", "purrr", "tidyr", "dplyr", "ggplot2", "ggtree", "hisse", "RColorBrewer", "readxl", "lattice", "writexl", "grid", "gridExtra", "gtable", "ggstance", "coda", "foreach", "cowplot", "geiger", "BlambdaV", "viridis", "nlme", "rr2")
lapply(pkg, require, character.only = TRUE)

```


## Load data

```{r}
## Load data

m_df <- readxl::read_excel("Tab_S2.xlsx", sheet = 2, na = c("","NA"),
                                 col_types = c("text", "text", rep("numeric",8)))


s_df <- readxl::read_excel("Tab_S3.xlsx", sheet = 2, na = c("","NA"),
                                  col_types = c(
                                    rep("text",5),
                                    rep("numeric",10),
                                    "text",
                                    rep("numeric",12),
                                    "text"
                                    )
                                 )

N_df <- readxl::read_excel("Tab_S1.xlsx", sheet = 2) %>%
  mutate(
    Species = gsub("M\\.", "Macrostomum", Species)
  )
```

## Transform the data

We want to do mean(log10(x)) *NOT* log(mean(x)). 
Since the former is a much better representation of a species value.


```{r}
morpho_df <- m_df %>%
  group_by(sci_name, ID) %>%
  mutate_at(c("body_size"), funs(body_size_logsqrt = log10(sqrt(.)))) %>%
  mutate_at(c("stylet_length", "stylet_proximal", "stylet_distal", "sperm_length", "bristle_length"),
            funs(log = log10(.))) %>%
  mutate_at(c( "bristle_length", "stylet_asymmetry"), funs(log = log10(. + 1)))

morpho_df2 <- morpho_df %>%
  group_by(sci_name) %>%
  select(sci_name, contains("log"), stylet_curvyness) %>%
  summarise_if(is.numeric, 
               funs(mean = mean(., na.rm = TRUE), 
                       N = sum(!is.na(.)))
               ) %>%
  mutate( sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", sci_name)))
```

## LOAD TREES

```{r}
trees <- readRDS(file = paste0("trees.Rdata"))
ltre <- length(trees)

# Color the hypodermic clade
t_injecting_clade <- extract.clade(trees[[3]],154)
names_injecting <- t_injecting_clade$tip.label

s_df$clade <- ifelse(s_df$Short %in% names_injecting, "clade1", "clade2")

species_df <- s_df %>%
  mutate(sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", Species)),
         inferred_syndrome_2states = case_when(
           inferred_syndrome == "intermediate" ~ "hypodermic",
           TRUE ~ inferred_syndrome)
         ) %>%
  select(-contains("mean"))

# Some in the species list have not been measured
species_df$Species %in% morpho_df2$sci_name
sum(species_df$Species %in% morpho_df2$sci_name)

# All measured are in the species list
morpho_df2$sci_name %in% species_df$Species

# Combine them
main_df <- left_join(species_df, morpho_df2, by = "sci_name_short")

length(species_df$Short)
species_df$Short %in% trees[[3]]$tip.label

```

## Color palette

```{r}
Nclusters=3
cl <- parallel::makeCluster(Nclusters)
doParallel::registerDoParallel(cl)
library(foreach)

pr = "#006837"
int = "#78c679"
ab = "#ffffcc"

palette1 <- c(pr, int, ab, "grey")
palette2 <- c(ab, int, pr, "grey")
palette3 <- c(ab, pr, "grey")
```


### multiply trees by 100 

The ML optimizer can run into issues due to numeric underflow. 

```{r}
trees <- lapply(trees,function(tre){tre$edge.length <- tre$edge.length * 100; return(tre)})
```

Make sure we are using factors

```{r}
main_df <- main_df %>%
  mutate(Rec_binary = factor(Rec_binary, levels = c(0,1)),
         bristle_binary = factor(bristle_binary, levels = c(0,1)),
         inferred_syndrome_2states = factor(inferred_syndrome_2states, 
                                            levels = c("hypodermic","reciprocal")))
```

# Prepare data for PCA

We want to use both the catgorical as well as the quantitative traits.
To make them comparable they need to be scaled. So all traits will be transformed
to be between 0 and 1. 

```{r}
# In previous version things had different names
main_df <- main_df %>%
  rename(
    brstlLen_um_per_sperm_log_mean = bristle_length_log_mean,
    bdSize_um_logsqrt_mean = body_size_logsqrt_mean, 
    sty_avg_um_log_mean = stylet_length_log_mean,
    wb_um_log_mean = stylet_proximal_log_mean,
    wt_um_log_mean = stylet_distal_log_mean,
    delta_e_um_log_mean = stylet_asymmetry_log_mean,
    total_sprmLen_um_log_mean = sperm_length_log_mean,
    sty_curveSum_avg_mean = stylet_curvyness_mean
  )
  

  
var_of_interest <- c("brstlLen_um_per_sperm_log_mean",
  "bdSize_um_logsqrt_mean", 
  "sty_avg_um_log_mean", "wb_um_log_mean", "wt_um_log_mean", "delta_e_um_log_mean",
  "total_sprmLen_um_log_mean","eye", "Rec", 
  "ant_thickness", "ant_cellular_valve", 
  "ant_chambers", "No_openings",
  "bristle", "brush", "velum",
   "sharp", "an_complex_noc_noo")

var_of_interest[!var_of_interest %in% names(main_df)]
```

# Prepare data for PCA

We want to use both the catgorical as well as the quantitative traits.
To make them comparable they need to be scaled. So all traits will be transformed
to be between 0 and 1. 

```{r}
var_of_interest <- c("brstlLen_um_per_sperm_log_mean",
  "bdSize_um_logsqrt_mean", 
  "sty_avg_um_log_mean", "wb_um_log_mean", "wt_um_log_mean", "delta_e_um_log_mean",
  "total_sprmLen_um_log_mean","eye", "Rec", 
  "ant_thickness", "ant_cellular_valve", 
  "ant_chambers", "No_openings",
  "bristle", "brush", "velum",
   "sharp", "an_complex_noc_noo")


quant_df <- main_df %>%
  select(Short,one_of(var_of_interest)) %>%
  rowwise() %>%
  as.data.frame()

scaled_df <- quant_df %>%
  mutate_if(is.numeric, funs(scales::rescale), to=c(0,1)) %>%
  rowwise() %>%
  as.data.frame()

row.names(quant_df) <- quant_df$Short
row.names(scaled_df) <- scaled_df$Short
```


Mark clade
```{r}
# This is dependent on this version of the 28S tree! 
t_injecting_clade <- extract.clade(trees[[3]],154)
names_injecting <- t_injecting_clade$tip.label

main_df$clade <- ifelse(main_df$Short %in% names_injecting, "clade1", "clade2")
```


Make sure we are using factors

```{r}
main_df <- main_df %>% 
  mutate(Rec_binary = factor(Rec_binary, levels = c(0,1)),
         bristle_binary = factor(bristle_binary, levels = c(0,1)),
         sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", Species)),
         inferred_syndrome_2states = case_when(
           inferred_syndrome == "intermediate" ~ "hypodermic",
           TRUE ~ inferred_syndrome
         )) %>%
  mutate(sharp = as.factor(sharp)) %>%
  mutate(taxa = sci_name_short) %>%
  as.data.frame()
```

# PCA 

## Antrum 

```{r}
i = 3
treename = "Short"
ppca_antrum <- main_df %>%
  select(Short,ant_thickness, ant_cellular_valve, ant_chambers, No_openings) %>%
  prepare_phyl_pca(tree = trees[[i]])

ignore = c("")
tmp_df <- ppca_antrum[[1]][,setdiff(names(ppca_antrum[[1]]), c(treename, ignore))]
t <- ppca_antrum[[2]]

pPCA_1 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
pPCA_1_L <- as.data.frame(pPCA_1$L) %>%
  mutate(var = rownames(pPCA_1$L))

pPCA_1_S <- as.data.frame(pPCA_1$S)  %>%
  mutate(Short = rownames(pPCA_1$S))

pPCA_p1 <- ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
  geom_point(size=4) 

summary(pPCA_1)
round(pPCA_1_L$PC1,2)
```

## Stylet

```{r}
ppca_stylet <- main_df %>%
    select(Short, sty_avg_um_log_mean,
           sty_curveSum_avg_mean, wb_um_log_mean, wt_um_log_mean,
           delta_e_um_log_mean) %>%
  prepare_phyl_pca(tree = trees[[i]])

ignore = c("")
tmp_df <- ppca_stylet[[1]][,setdiff(names(ppca_stylet[[1]]), c(treename, ignore))]
t <- ppca_stylet[[2]]

pPCA_2 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
pPCA_2_L <- as.data.frame(pPCA_2$L) %>%
  mutate(var = rownames(pPCA_2$L))

pPCA_2_S <- as.data.frame(pPCA_2$S)  %>%
  mutate(Short = rownames(pPCA_2$S))

pPCA_p2 <- ggplot(as.data.frame(pPCA_2$S), aes(x = PC1, y=PC2)) +
  geom_point(size=4)

summary(pPCA_2)
round(pPCA_2_L$PC1,2)
```

Join the two PCA results

```{r}
pPCA_1
pPCA_2

names(pPCA_1_S) <- c("anPC1", "anPC2", "anPC3", "anPC4", "Short")
# reverse them to make them easier to think about
pPCA_1_S$anPC1 <- pPCA_1_S$anPC1 * -1 


names(pPCA_2_S) <- c("styPC1", "styPC2", "styPC3", "styPC4", "styPC5", "Short")
# pPCA_2_S$styPC1 <- pPCA_2_S$styPC1 * -1 

PCA_df <- left_join(pPCA_1_S, pPCA_2_S) %>%
  na.omit() %>%
  left_join(select(main_df, Short, inferred_syndrome,inferred_syndrome_2states))

ggplot(PCA_df, aes(x = anPC1, y=styPC1, fill=inferred_syndrome)) +
    geom_point(size=4, pch=21) 
```

# PGLS test of PCA first axis

## Across all species

```{r}
formula = "styPC1 ~ anPC1"
N_var = ""
predictor <- "styPC1"
pred <- "anPC1"
treename = "Short"
cor_form = "~ Short"
DIAGNOSTIC = FALSE

pgls_styPC1_anPC1_all <- list()
pgls_styPC1_anPC1_all_models <- list()
for (j in 1:3){
  # Calculate pPCA antrum
  ppca_antrum <- main_df %>%
  select(Short,ant_thickness, ant_cellular_valve, ant_chambers, No_openings) %>%
  prepare_phyl_pca(tree = trees[[j]])
  ignore = c("")
  tmp_df <- ppca_antrum[[1]][,setdiff(names(ppca_antrum[[1]]), c(treename, ignore))]
  t <- ppca_antrum[[2]]
  pPCA_1 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
  pPCA_1_L <- as.data.frame(pPCA_1$L) %>%
    mutate(var = rownames(pPCA_1$L))

  pPCA_1_S <- as.data.frame(pPCA_1$S)  %>%
    mutate(Short = rownames(pPCA_1$S))
  
  # Calculate pPCA stylet 
  ppca_stylet <- main_df %>%
    select(Short, sty_avg_um_log_mean,
           sty_curveSum_avg_mean, wb_um_log_mean, wt_um_log_mean,
           delta_e_um_log_mean) %>%
    prepare_phyl_pca(tree = trees[[j]])
  ignore = c("")
  tmp_df <- ppca_stylet[[1]][,setdiff(names(ppca_stylet[[1]]), c(treename, ignore))]
  t <- ppca_stylet[[2]]
  

  pPCA_2 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
  pPCA_2_L <- as.data.frame(pPCA_2$L) %>%
    mutate(var = rownames(pPCA_2$L))

  pPCA_2_S <- as.data.frame(pPCA_2$S)  %>%
    mutate(Short = rownames(pPCA_2$S))
  
  # join PCA dfs.
  names(pPCA_1_S) <- c("anPC1", "anPC2", "anPC3", "anPC4", "Short")
  names(pPCA_2_S) <- c("styPC1", "styPC2", "styPC3", "styPC4", "styPC5", "Short")
    # reverse them to make them easier to think about
  pPCA_1_S$anPC1 <- pPCA_1_S$anPC1 * -1 
  pPCA_1_S$anlambda <- round(pPCA_1$lambda, 2)
  pPCA_1_L$PC1 <- pPCA_1_L$PC1 * -1 
  #pPCA_2_S$styPC1 <- pPCA_2_S$styPC1 * -1 
  pPCA_2_S$stylambda <- round(pPCA_2$lambda, 2)
  #pPCA_2_L$PC1 <- pPCA_2_L$PC1 * -1 
  
  PCA_df <- left_join(pPCA_1_S, pPCA_2_S) %>%
    na.omit() %>%
    left_join(select(main_df, Short, sci_name_short, inferred_syndrome, inferred_syndrome_2states))
  
  write.csv(PCA_df, file=paste0("output/EL_PCA_df_", names(trees)[[3]],".csv"))
  # Output loadings and scores
  write.csv(pPCA_1_L, file=paste0("output/EL_anPC_L_", names(trees)[[j]],".csv"), row.names = FALSE)
  
  pPCA_1_S %>%
    left_join(select(main_df, Short, sci_name_short)) %>%
    #mutate_if(is.numeric, round, 2) %>%
    select(sci_name_short, contains("PC"), anlambda) %>%
    rename(lambda = anlambda) %>%
    write.csv(file=paste0("output/EL_anPC_S_", names(trees)[[j]],".csv"), row.names = FALSE)

  write.csv(pPCA_2_L, file=paste0("output/EL_styPC_L_", names(trees)[[j]],".csv"), row.names = FALSE)
  
  pPCA_2_S %>%
    left_join(select(main_df, Short, sci_name_short)) %>%
    #mutate_if(is.numeric, round, 2) %>%
    select(sci_name_short, contains("PC"), stylambda) %>%
    rename(lambda = stylambda) %>%
    write.csv(file=paste0("output/EL_styPC_S_", names(trees)[[j]],".csv"), row.names = FALSE)

  ggplot(PCA_df, aes(x = anPC1, y=styPC1, fill=inferred_syndrome)) +
  geom_point(size=4, pch=21) 
  
  # PGLS
  comp_df <- PCA_df %>%
    select(Short, pred, predictor, inferred_syndrome) %>%
  prepare_phyl_pca(tree = trees[[j]], treename = "Short")
  
  # Script for model testing is omitted here since the lambda model was always preferred.
  # See PGLS.Rmd for details.
  
  lambda <- tryCatch(
    {gls(as.formula(formula), data = comp_df[[1]], 
                 correlation = corPagel(1, phy=comp_df[[2]], form = as.formula(cor_form)), 
                  method = "REML",
                control=list(maxIter=500))},
    error=function(cond){
      print(cond)
      optimal_lambda <- optimize_lambda(
        formula=as.formula(formula), data = comp_df[[1]], phy=comp_df[[2]], form = as.formula(cor_form),
         method = "REML", control=list(), weight = NULL)
      gls(as.formula(formula), data = comp_df[[1]], 
                 correlation = corPagel(optimal_lambda, 
                                        phy=comp_df[[2]], fixed=TRUE, form = as.formula(cor_form)), 
                 method = "REML",
                 control=list(maxIter=500))})


  ############
  mod0 <- gls(as.formula("styPC1 ~ 1"), data = comp_df[[1]], 
                 correlation = corPagel(1,
                                        phy=comp_df[[2]], form = ~Short),
                 method = "REML",
                control=list(maxIter=500))
  
  full =   lambda 
  phy_only = mod0
  data_only = lm(as.formula(formula), data = comp_df[[1]])
  R2_pred_data = round(R2.pred(full, phy_only),3)
  R2_pred_phy = round(R2.pred(full, data_only),3)
  R2_pred_full = round(R2.pred(full),3)
  
  #####################
  s_best <- summary(lambda)
  res_df <- data.frame(s_best$tTable)
  res_df$Estimate <- rownames(s_best$tTable)
  res_df$model <- "lambda"
  res <- res_df
  res$R2pred <- c(R2_pred_full, rep(R2_pred_data, nrow(res)-1))
  
  res$predictor = pred
  res$phylogeny = names(trees)[[j]]
  res$lambda = lambda$modelStruct[[1]]
  res$N = nrow(comp_df[[1]])
  #############
  # DIAGNOSTICS
  if (DIAGNOSTIC == TRUE) {
    png(paste0("fig/diag/",outname, "_", names(trees)[j],"_",preds[i],".png"),
      width = 7, height = 7, unit="in", res = 300)
    pgls_diag_gls(name=paste0(names(trees)[j]," ",predictor), mod_type = "lambda",
                model=lambda, formula=as.formula(formulas[[i]]),
                data = comp_df[[1]], phy=comp_df[[2]], 
                weight = as.formula(N_var), method = "REML", control=list(maxIter=500, opt="optim"))
    dev.off()
  }

  ############
  pgls_styPC1_anPC1_all <- append(pgls_styPC1_anPC1_all,list(res))
  pgls_styPC1_anPC1_all_models <- append(pgls_styPC1_anPC1_all_models, list(lambda))
}



main_res <- do.call(rbind, pgls_styPC1_anPC1_all)
main_res$p.value <- format.pval(main_res$p.value, eps=0.001, digits=2)
main_res_out <- main_res %>%
  select(predictor, phylogeny, Estimate, N, lambda, Value, Std.Error, t.value, p.value, R2pred) %>%
  mutate_if(is.numeric, round, 2)



write.csv(main_res_out, file="output/EL_styPC1_anPC1_all_PGLS.csv")

names(pgls_styPC1_anPC1_all_models) <- apply(expand.grid(names(trees)[1:3], pred), 1, paste, collapse="_")

saveRDS(pgls_styPC1_anPC1_all_models, "output/EL_styPC1_anPC1_all_PGLS.RDS")
```

## Reciprocal species only

```{r}
formula = "styPC1 ~ anPC1"
N_var = ""
predictor <- "styPC1"
pred <- "anPC1"
treename = "Short"
cor_form = "~ Short"
DIAGNOSTIC = FALSE

pgls_styPC1_anPC1_reciprocal <- list()
pgls_styPC1_anPC1_reciprocal_models <- list()
for (j in 1:3){
  # Calculate pPCA antrum
  ppca_antrum <- main_df %>%
  select(Short,ant_thickness, ant_cellular_valve, ant_chambers, No_openings) %>%
  prepare_phyl_pca(tree = trees[[j]])
  ignore = c("")
  tmp_df <- ppca_antrum[[1]][,setdiff(names(ppca_antrum[[1]]), c(treename, ignore))]
  t <- ppca_antrum[[2]]
  pPCA_1 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
  pPCA_1_L <- as.data.frame(pPCA_1$L) %>%
    mutate(var = rownames(pPCA_1$L))

  pPCA_1_S <- as.data.frame(pPCA_1$S)  %>%
    mutate(Short = rownames(pPCA_1$S))
  
  # Calculate pPCA stylet 
  ppca_stylet <- main_df %>%
    select(Short, sty_avg_um_log_mean,
           sty_curveSum_avg_mean, wb_um_log_mean, wt_um_log_mean,
           delta_e_um_log_mean) %>%
    prepare_phyl_pca(tree = trees[[j]])
  ignore = c("")
  tmp_df <- ppca_stylet[[1]][,setdiff(names(ppca_stylet[[1]]), c(treename, ignore))]
  t <- ppca_stylet[[2]]
  

  pPCA_2 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
  pPCA_2_L <- as.data.frame(pPCA_2$L) %>%
    mutate(var = rownames(pPCA_2$L))

  pPCA_2_S <- as.data.frame(pPCA_2$S)  %>%
    mutate(Short = rownames(pPCA_2$S))
  
  # join PCA dfs.
  names(pPCA_1_S) <- c("anPC1", "anPC2", "anPC3", "anPC4", "Short")
  names(pPCA_2_S) <- c("styPC1", "styPC2", "styPC3", "styPC4", "styPC5", "Short")
    # reverse them to make them easier to think about
  pPCA_1_S$anPC1 <- pPCA_1_S$anPC1 * -1 
  pPCA_1_S$anlambda <- round(pPCA_1$lambda, 2)
  pPCA_1_L$PC1 <- pPCA_1_L$PC1 * -1 
  #pPCA_2_S$styPC1 <- pPCA_2_S$styPC1 * -1 
  pPCA_2_S$stylambda <- round(pPCA_2$lambda, 2)
  #pPCA_2_L$PC1 <- pPCA_2_L$PC1 * -1 
  
  PCA_df <- left_join(pPCA_1_S, pPCA_2_S) %>%
    na.omit() %>%
    left_join(select(main_df, Short, sci_name_short, inferred_syndrome, inferred_syndrome_2states))
  
  # Output loadings and scores
  # write.csv(pPCA_1_L, file=paste0("output/11_anPC_L_", names(trees)[[j]],".csv"), row.names = FALSE)
  # 
  # pPCA_1_S %>%
  #   left_join(select(main_df, Short, sci_name_short)) %>%
  #   #mutate_if(is.numeric, round, 2) %>%
  #   select(sci_name_short, contains("PC"), anlambda) %>%
  #   rename(lambda = anlambda) %>%
  #   write.csv(file=paste0("output/EL_anPC_S_", names(trees)[[j]],".csv"), row.names = FALSE)
  # 
  # write.csv(pPCA_2_L, file=paste0("output/EL_styPC_L_", names(trees)[[j]],".csv"), row.names = FALSE)
  # 
  # pPCA_2_S %>%
  #   left_join(select(main_df, Short, sci_name_short)) %>%
  #   #mutate_if(is.numeric, round, 2) %>%
  #   select(sci_name_short, contains("PC"), stylambda) %>%
  #   rename(lambda = stylambda) %>%
  #   write.csv(file=paste0("output/EL_styPC_S_", names(trees)[[j]],".csv"), row.names = FALSE)

  # ggplot(PCA_df, aes(x = anPC1, y=styPC1, fill=inferred_syndrome)) +
  # geom_point(size=4, pch=21) 
  # 
  # PGLS
  comp_df <- PCA_df %>%
    filter(inferred_syndrome == "reciprocal") %>%
    select(Short, pred, predictor, inferred_syndrome) %>%
  prepare_phyl_pca(tree = trees[[j]], treename = "Short")
  
  # Script for model testing is omitted here since the lambda model was always preferred.
  # See PGLS.Rmd for details.
  
  lambda <- tryCatch(
    {gls(as.formula(formula), data = comp_df[[1]], 
                 correlation = corPagel(1, phy=comp_df[[2]], form = as.formula(cor_form)), 
                  method = "REML",
                control=list(maxIter=500))},
    error=function(cond){
      print(cond)
      optimal_lambda <- optimize_lambda(
        formula=as.formula(formula), data = comp_df[[1]], phy=comp_df[[2]], form = as.formula(cor_form),
         method = "REML", control=list(), weight = NULL)
      gls(as.formula(formula), data = comp_df[[1]], 
                 correlation = corPagel(optimal_lambda, 
                                        phy=comp_df[[2]], fixed=TRUE, form = as.formula(cor_form)), 
                 method = "REML",
                 control=list(maxIter=500))})


  ############
  mod0 <- gls(as.formula("styPC1 ~ 1"), data = comp_df[[1]], 
                 correlation = corPagel(1,
                                        phy=comp_df[[2]], form = ~Short),
                 method = "REML",
                control=list(maxIter=500))
  
  full =   lambda 
  phy_only = mod0
  data_only = lm(as.formula(formula), data = comp_df[[1]])
  R2_pred_data = round(R2.pred(full, phy_only),3)
  R2_pred_phy = round(R2.pred(full, data_only),3)
  R2_pred_full = round(R2.pred(full),3)
  
  #####################
  s_best <- summary(lambda)
  res_df <- data.frame(s_best$tTable)
  res_df$Estimate <- rownames(s_best$tTable)
  res_df$model <- "lambda"
  res <- res_df
  res$R2pred <- c(R2_pred_full, rep(R2_pred_data, nrow(res)-1))
  
  res$predictor = pred
  res$phylogeny = names(trees)[[j]]
  res$lambda = lambda$modelStruct[[1]]
  res$N = nrow(comp_df[[1]])
  #############
  # DIAGNOSTICS
  if (DIAGNOSTIC == TRUE) {
    png(paste0("fig/diag/",outname, "_", names(trees)[j],"_",preds[i],".png"),
      width = 7, height = 7, unit="in", res = 300)
    pgls_diag_gls(name=paste0(names(trees)[j]," ",predictor), mod_type = "lambda",
                model=lambda, formula=as.formula(formulas[[i]]),
                data = comp_df[[1]], phy=comp_df[[2]], 
                weight = as.formula(N_var), method = "REML", control=list(maxIter=500, opt="optim"))
    dev.off()
  }

  ############
  pgls_styPC1_anPC1_reciprocal <- append(pgls_styPC1_anPC1_reciprocal,list(res))
  pgls_styPC1_anPC1_reciprocal_models <- append(pgls_styPC1_anPC1_reciprocal_models, list(lambda))
}



main_res <- do.call(rbind, pgls_styPC1_anPC1_reciprocal)
main_res$p.value <- format.pval(main_res$p.value, eps=0.001, digits=2)
main_res_out <- main_res %>%
  select(predictor, phylogeny, Estimate, N, lambda, Value, Std.Error, t.value, p.value, R2pred) %>%
  mutate_if(is.numeric, round, 2)


write.csv(main_res_out, file="output/EL_styPC1_anPC1_reciprocal_PGLS.csv")

names(pgls_styPC1_anPC1_reciprocal_models) <- apply(expand.grid(names(trees)[1:3], pred), 1, paste, collapse="_")

saveRDS(pgls_styPC1_anPC1_reciprocal_models, "output/EL_styPC1_anPC1_reciprocal_PGLS.RDS")
```

# Plot for the C-IQ-TREE results

Load models

```{r}
pPCA_all_models   <- readRDS("output/EL_styPC1_anPC1_all_PGLS.RDS")
pPCA_reciprocal_models <- readRDS("output/EL_styPC1_anPC1_reciprocal_PGLS.RDS")
PCA_df <- read.csv(file=paste0("output/EL_PCA_df_", names(trees)[[3]],".csv"))

an_PCA_L <- read.csv(file = "output/EL_anPC_L_C-IQ-TREE.csv")
sty_PCA_L <- read.csv(file = "output/EL_styPC_L_C-IQ-TREE.csv")

```

Show loadings
```{r}

an_heatp <-   an_PCA_L %>%
  gather(key = PC, value = val, PC1) %>%
  mutate(var = factor(var, levels = rev(unique(an_PCA_L$var)))) %>%
  ggplot(aes(x = PC, y = var)) +
  geom_tile(aes(fill=val)) +
  geom_text(aes(label=round(val,2)), size =6) + 
  scale_fill_gradient2(low="royalblue", mid="white", high="firebrick1", limits=c(-1,1)) +
  scale_y_discrete(labels = 
  c('No. genital pores','Chamber complexity','Cellular valve', 'Thickness')) +
  scale_x_discrete(labels = "PC1 (46.9%)") +
  theme(legend.position = "none",
        axis.title = element_blank())

sty_heatp <-   sty_PCA_L %>%
  gather(key = PC, value = val, PC1) %>%
  mutate(var = factor(var, levels = rev(unique(sty_PCA_L$var)))) %>%
  ggplot(aes(x = PC, y = var)) +
  geom_tile(aes(fill=val)) +
  geom_text(aes(label=round(val,2)), size =6) + 
  scale_fill_gradient2(low="royalblue", mid="white", high="firebrick1", limits=c(-1,1)) +
  scale_y_discrete(labels = 
  c('Distal asymmetry','Width distal','Width proximal', 'Curviness', 'Stylet length')) +
  scale_x_discrete(labels = "PC1 (38.6%)") +
  theme(legend.position = "none",
        axis.title = element_blank())
p_loadings <- plot_grid(sty_heatp, an_heatp, ncol = 1, align = "v")
  
```

# Plotting the results

```{r}
formula = "styPC1 ~ anPC1"
N_var = ""
predictor <- "styPC1"
pred <- "anPC1"
treename = "Short"


preds <- list("styPC1")
j=3
pdf(file = "fig/EL_PGLS_pPCA_anPC1_styPC1.pdf", useDingbats = FALSE)

comp_df <- list()
comp_df[[1]] <- PCA_df 
# parameters to plot slopes
mod <- pPCA_all_models[["C-IQ-TREE_anPC1"]]
inter <- mod$coefficients[[1]]; slope <- mod$coefficients[[2]]
xlow <-  min(comp_df[[1]][pred])
xhigh <-  max(comp_df[[1]][pred])
ylow <- inter + xlow * slope; yhigh <- inter + xhigh * slope

# 
R_mod <- pPCA_reciprocal_models[["C-IQ-TREE_anPC1"]]
R_inter <- R_mod$coefficients[[1]]; R_slope <- R_mod$coefficients[[2]]
R_xlow <-  min(
  filter(comp_df[[1]], inferred_syndrome == "reciprocal")[pred])
R_xhigh <-  max(
  filter(comp_df[[1]], inferred_syndrome == "reciprocal")[pred])
R_ylow <- R_inter + R_xlow * R_slope; R_yhigh <- R_inter + R_xhigh * R_slope

gplot <- ggplot(comp_df[[1]], aes_string(x=pred, y="styPC1",
                         fill = "inferred_syndrome")) +
  geom_segment(x = xlow, xend = xhigh, y = ylow, yend = yhigh, size=1) +
  geom_segment(x = R_xlow, xend = R_xhigh, y = R_ylow, yend = R_yhigh,
               linetype = "dashed", size=1) +
  geom_point(size=3, shape=21, stroke=1) +
  scale_fill_manual(values=palette2) +
  scale_shape_manual(values=c(21,24,22,21)) +
  ylab("Stylet PC1") +
  xlab("Antrum PC1") +
  theme_cowplot() +
  theme(legend.position=c(.05, 0.9),
        legend.title = element_blank())

dev.off()

```


```{r}
pdf("fig/EL_sty_an_PCA_loadings.pdf", width = 10, height = 7)
plot_grid(p_loadings,
          gplot, rel_widths = c(3,7), align = "h")
dev.off()
```

