---
title: "08_PCA"
author: "Jeremias Brand"
date: "28 Mai 2019"
output: html_document
---

```{r}
require(knitr)
opts_knit$set(root.dir = "..")
```



```{r results="hide", warning=FALSE }

source("R/functions.R")
library(dplyr)
library(scales)
library(tidyr)
library(ggplot2)
library(readxl)
library(purrr)
library(ape)
library(phytools)
library(geiger)
library(phangorn)
library(forcats)
library(viridis)
library(ggtree)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(gtable)
library(stringr)
library(ICC)
library(lmtest)
library(nlme)
library(ggrepel)
library(cowplot)
library(rr2)
library(animation)

cols <- c("black", "lightblue", "blue", "Grey")
shapes <- c(19,17,15,1)
treename = "Short"


pr = "#006837"
int = "#78c679"
ab = "#ffffcc"

palette1 <- c(pr, int, ab, "grey")
palette2 <- c(ab, int, pr, "grey")
palette3 <- c(ab, pr, "grey")
# Prevent MASS and dplyr clashes
select <- dplyr::select
```

## Load data

```{r}
## Load data

m_df <- readxl::read_excel("Tab_S2.xlsx", sheet = 2, na = c("","NA"),
                                 col_types = c("text", "text", rep("numeric",8)))


s_df <- readxl::read_excel("Tab_S3.xlsx", sheet = 2, na = c("","NA"),
                                  col_types = c(
                                    rep("text",5),
                                    rep("numeric",10),
                                    "text",
                                    rep("numeric",12),
                                    "text"
                                    )
                                 )

N_df <- readxl::read_excel("Tab_S1.xlsx", sheet = 2) %>%
  mutate(
    Species = gsub("M\\.", "Macrostomum", Species)
  )
```

## Transform the data

We want to do mean(log10(x)) *NOT* log(mean(x)). 
Since the former is a much better representation of a species value.


```{r}
morpho_df <- m_df %>%
  group_by(sci_name, ID) %>%
  mutate_at(c("body_size"), funs(body_size_logsqrt = log10(sqrt(.)))) %>%
  mutate_at(c("stylet_length", "stylet_proximal", "stylet_distal", "sperm_length", "bristle_length"),
            funs(log = log10(.))) %>%
  mutate_at(c( "bristle_length", "stylet_asymmetry"), funs(log = log10(. + 1)))

morpho_df2 <- morpho_df %>%
  group_by(sci_name) %>%
  select(sci_name, contains("log"), stylet_curvyness) %>%
  summarise_if(is.numeric, 
               funs(mean = mean(., na.rm = TRUE), 
                       N = sum(!is.na(.)))
               ) %>%
  mutate( sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", sci_name)))
```

## LOAD TREES

```{r}
trees <- readRDS(file = paste0("trees.Rdata"))
ltre <- length(trees)

# Color the hypodermic clade
t_injecting_clade <- extract.clade(trees[[3]],154)
names_injecting <- t_injecting_clade$tip.label

s_df$clade <- ifelse(s_df$Short %in% names_injecting, "clade1", "clade2")

species_df <- s_df %>%
  mutate(sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", Species)),
         inferred_syndrome_2states = case_when(
           inferred_syndrome == "intermediate" ~ "hypodermic",
           TRUE ~ inferred_syndrome)
         ) %>%
  select(-contains("mean"))

# Some in the species list have not been measured
species_df$Species %in% morpho_df2$sci_name
sum(species_df$Species %in% morpho_df2$sci_name)

# All measured are in the species list
morpho_df2$sci_name %in% species_df$Species

# Combine them
main_df <- left_join(species_df, morpho_df2, by = "sci_name_short")

length(species_df$Short)
species_df$Short %in% trees[[3]]$tip.label

```

## Color palette

```{r}
Nclusters=3
cl <- parallel::makeCluster(Nclusters)
doParallel::registerDoParallel(cl)
library(foreach)

pr = "#006837"
int = "#78c679"
ab = "#ffffcc"

palette1 <- c(pr, int, ab, "grey")
palette2 <- c(ab, int, pr, "grey")
palette3 <- c(ab, pr, "grey")
```


### multiply trees by 100 

The ML optimizer can run into issues due to numeric underflow. 

```{r}
trees <- lapply(trees,function(tre){tre$edge.length <- tre$edge.length * 100; return(tre)})
```

Make sure we are using factors

```{r}
main_df <- main_df %>%
  mutate(Rec_binary = factor(Rec_binary, levels = c(0,1)),
         bristle_binary = factor(bristle_binary, levels = c(0,1)),
         inferred_syndrome_2states = factor(inferred_syndrome_2states, 
                                            levels = c("hypodermic","reciprocal")))
```

# Prepare data for PCA

We want to use both the catgorical as well as the quantitative traits.
To make them comparable they need to be scaled. So all traits will be transformed
to be between 0 and 1. 

```{r}
# In previous version things had different names
main_df <- main_df %>%
  rename(
    brstlLen_um_per_sperm_log_mean = bristle_length_log_mean,
    bdSize_um_logsqrt_mean = body_size_logsqrt_mean, 
    sty_avg_um_log_mean = stylet_length_log_mean,
    wb_um_log_mean = stylet_proximal_log_mean,
    wt_um_log_mean = stylet_distal_log_mean,
    delta_e_um_log_mean = stylet_asymmetry_log_mean,
    total_sprmLen_um_log_mean = sperm_length_log_mean,
    sty_curveSum_avg_mean = stylet_curvyness_mean
  )
  
var_of_interest[!var_of_interest %in% names(main_df)]
  
var_of_interest <- c("brstlLen_um_per_sperm_log_mean",
  "bdSize_um_logsqrt_mean", 
  "sty_avg_um_log_mean", "wb_um_log_mean", "wt_um_log_mean", "delta_e_um_log_mean",
  "total_sprmLen_um_log_mean","eye", "Rec", 
  "ant_thickness", "ant_cellular_valve", 
  "ant_chambers", "No_openings",
  "bristle", "brush", "velum",
   "sharp", "an_complex_noc_noo")


quant_df <- main_df %>%
  select(Short,one_of(var_of_interest)) %>%
  rowwise() %>%
  as.data.frame()

scaled_df <- quant_df %>%
  mutate_if(is.numeric, funs(scales::rescale), to=c(0,1)) %>%
  rowwise() %>%
  as.data.frame()

row.names(quant_df) <- quant_df$Short
row.names(scaled_df) <- scaled_df$Short
```

## Plot all traits next to tree

```{r}
for (i in 1:ltre) {
pdf(paste0("fig/scaled_traits_plotted_",names(trees)[[i]],".pdf"), height = 24, width = 10)
t_UM = trees[[i]]
col_pal <- viridis(n=10, option = "A")
col_l = col_pal[[1]]
col_h = col_pal[[10]]
p <- ggtree(t_UM) + geom_tiplab(size=2)
print(gheatmap(p, select(scaled_df, -(Short:Rec)), offset=0.05, width=0.2,
         low=col_l, high=col_h, colnames_position = "top",
         colnames_angle = 90, colnames_offset_y = 2.6,  font.size=2) )

print(gheatmap(p, select(scaled_df, (ant_thickness:bristle)), offset=0.05, width=0.2,
         low=col_l, high=col_h, colnames_position = "top",
         colnames_angle = 90, colnames_offset_y = 2.6,  font.size=2)) 

print(gheatmap(p, select(scaled_df, c(an_complex_noc_noo,bristle)), offset=0.05, width=0.2,
         low=col_l, high=col_h, colnames_position = "top",
         colnames_angle = 90, colnames_offset_y = 2.6,  font.size=2))

print(gheatmap(p, select(scaled_df, Rec, bristle), offset=0.05, width=0.2,
         low=col_l, high=col_h, colnames_position = "top",
         colnames_angle = 90, colnames_offset_y = 2.6,  font.size=2))
dev.off()
}
```


## PCA of all traits

```{r}
for (i in 1:ltre) {
  ignore = c("Rec", "bristle", "clade", "inferred_syndrome")
  
  ppca1 <- main_df %>%
    select(Short, Rec, bristle, clade, inferred_syndrome,
           # stylet traits
           sty_avg_um_log_mean,
           wt_um_log_mean,
           wb_um_log_mean,
           sty_curveSum_avg_mean,
           #sty_curveSum_scaled_mean,
           delta_e_um_log_mean,
           sharp,
           # sperm traits
           total_sprmLen_um_log_mean,
           brstlLen_um_per_sperm_log_mean,
           velum,
           brush,
           # antrum traits
           ant_thickness,
           ant_cellular_valve,
           ant_chambers,
           No_openings,
           # body size
           bdSize_um_logsqrt_mean) %>%
    mutate(
      brush = ifelse(brush %in% c(1,2),1, ifelse(!is.na(brush),0,NA))
      ) %>% 
    rename(
      'stylet length (logum)' = sty_avg_um_log_mean,
      'width distal stylet opening (logum)' = wt_um_log_mean,
      'width proximal stylet opening (logum)' = wb_um_log_mean,
      'curviness of stylet' = sty_curveSum_avg_mean,
      #'curviness of stylet' = sty_curveSum_scaled_mean,
      'asymetry of distal thickening (logum)' = delta_e_um_log_mean,
      'sharpness of stylet' = sharp,
      # sperm traits
      'sperm length (logum)'=total_sprmLen_um_log_mean,
      'sperm bristle length (logum)' = brstlLen_um_per_sperm_log_mean,
      'sperm velum' = velum,
      'sperm brush' = brush,
      # antrum traits
      'antrum thickness' = ant_thickness,
      'antrum cellular valve' = ant_cellular_valve,
      'complexity of antrum chamber' = ant_chambers,
      'number of vaginas' = No_openings,
           # body size
      'body area (logsqrtum)' = bdSize_um_logsqrt_mean
      )%>% 
  as.data.frame() %>%
  prepare_phyl_pca(tree = trees[[i]], ignore=ignore)
  
  var_order <- c('stylet length (logum)',
      'curviness of stylet',
      'width proximal stylet opening (logum)',
      'width distal stylet opening (logum)',
      'asymetry of distal thickening (logum)',
      'sharpness of stylet',
      'sperm length (logum)',
      'sperm bristle length (logum)',
      'sperm velum',
      'sperm brush',
      'antrum thickness',
      'antrum cellular valve',
      'complexity of antrum chamber',
      'number of vaginas',
      'body area (logsqrtum)')
  # remove the ignored columns which are just used to paint later
  tmp_df <- ppca1[[1]][,setdiff(names(ppca1[[1]]), c(treename, ignore))]
  t <- ppca1[[2]]
  
  cols <- c("black", "#1FBDCC", "#FF0E98", "grey")
  
  pPCA_1 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")
  
    # I want the majority of the loadings to be positive
  if (sum(-1*pPCA_1$L[,1]) > sum(pPCA_1$L[,1])){
    pPCA_1$L[,1] <- -1*pPCA_1$L[,1]
    pPCA_1$S[,1] <- -1*pPCA_1$S[,1]
  }

  if (sum(-1*pPCA_1$L[,2]) > sum(pPCA_1$L[,2])){
    pPCA_1$L[,2] <- -1*pPCA_1$L[,2]
    pPCA_1$S[,2] <- -1*pPCA_1$S[,2]
  }
  
  pPCA_1_L <- as.data.frame(pPCA_1$L) %>%
  mutate(var = rownames(pPCA_1$L))
  
  pPCA_1_S <- as.data.frame(pPCA_1$S)  %>%
  mutate(Short = rownames(pPCA_1$S))
  
  ppca1[[1]]$Rec <- ifelse(is.na(ppca1[[1]]$Rec), "no_observation", ppca1[[1]]$Rec)
  
  main_df %>% 
    select(Short,inferred_syndrome, bristle, sharp,
         an_binary, an_complex_noc_noo, Rec) %>% 
    write.csv("output/inferred_syndrome_summary.csv")
  
  ## morphospace plot
  # color by label
  ppca1[[1]]$inferred_syndrome[match(ppca1[[2]]$tip.label, ppca1[[1]]$Short)]
  colmap <- setNames(palette2[1:length(unique(ppca1[[1]]$inferred_syndrome))],
                     sort(unique(ppca1[[1]]$inferred_syndrome)))
  ordered_trait <- ppca1[[1]]$inferred_syndrome[match(ppca1[[2]]$tip.label,
                                                      ppca1[[1]]$Short)]
  
  tipcols <- colmap[match(ordered_trait, names(colmap))]
  names(tipcols) <- ppca1[[2]]$tip.label
  t_cols<-c(tipcols[ppca1[[2]]$tip.label],rep("black",ppca1[[2]]$Nnode))
  names(t_cols)<-1:(length(ppca1[[2]]$tip)+ppca1[[2]]$Nnode)

  # ani.record(reset = TRUE)
  # project.phylomorphospace_mod(ppca1[[2]], pPCA_1$S[,1:2],
  #                          direction = "to", sleep=0, 
  #                          control=list(col.node=t_cols 
  #                                       ), node.size=c(0.8,1.8),
  #                          xlab="PC1",ylab="PC2")
  #   
  # oopts = ani.options(interval = 0.01)
  # ani.replay()
  # saveGIF(ani.replay(), img.name = "record_plot",
  #           ani.height = 1000, ani.width=1000)


  
  pPCA_1$L
  #####################
  ## 
  ####
  size = 8
  stroke = 1
  alpha = 0.8
  shapes <- c(21,24,22,21)
  #shapes = c(16,17,15,18)
 # cols <- c("black", "#1FBDCC", "#FF0E98", "grey")
  
  pPCA_plane <- ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
    geom_point(aes(shape= factor(ppca1[[1]]$bristle)),
               size=size, stroke = stroke,fill="black") +
    scale_shape_manual(labels = c("absent", "reduced", "present"), values = shapes) +
    labs(shape = "sperm bristle category", fill = "clade") +
    #theme(legend.position = "none")
        theme(legend.position = c(0.05, 0.8), legend.direction="horizontal") +
      guides(fill=guide_legend(override.aes=list(shape=21)))
  
  legend_pPCA_plane <- get_legend(pPCA_plane)
  pPCA_plane 
    
  pPCA_clade <- ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
    geom_point(aes(fill= factor(ppca1[[1]]$clade),
                   shape= factor(ppca1[[1]]$bristle)),
               size=size, stroke = stroke, alpha=alpha) +
    scale_fill_manual(labels = c("clade1", "clade2"), values = cols) +
    scale_shape_manual(labels = c("absent", "reduced", "present"), values = shapes) +
    labs(shape = "sperm bristle category", fill = "clade") +
    #theme(legend.position = "none")
    theme(legend.position = c(0.05, 0.8), legend.direction="horizontal") +
      guides(fill=guide_legend(override.aes=list(shape=21)))
  legend_pPCA_clade <- get_legend(pPCA_clade)
  
  pPCA_clade
  # reverse colors for consistentcy
  cols <- palette1
  pPCA_p1 <- ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
    geom_point(aes(shape= factor(ppca1[[1]]$bristle), fill=factor(ppca1[[1]]$Rec)),
               size=size, stroke = stroke, alpha=alpha) +
    scale_fill_manual(labels = c("an", "both", "hy", "no observation"), values = cols) + 
    scale_shape_manual(labels = c("absent", "reduced", "present"), values = shapes) + 
    labs(shape = "sperm bristle category", fill = "Received sperm location") +
    #theme(legend.position = "none")
    theme(legend.position = c(0.05, 0.8), legend.direction="horizontal") +
      guides(fill=guide_legend(override.aes=list(shape=21)))
  legend_pPCA_p1 <- get_legend(pPCA_p1)
  pPCA_p1 
  
  cols <- palette1
  pPCA_inferred <- ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
    geom_point(aes(shape= factor(ppca1[[1]]$bristle), fill=factor(ppca1[[1]]$inferred_syndrome)), 
               size=size, stroke = stroke, alpha=alpha) +
    scale_fill_manual(labels = c("reciprocal", "intermediate", "hypodermic", "no observation"), values = palette2) + 
    scale_shape_manual(labels = c("absent", "reduced", "present"), values = shapes) + 
    labs(shape = "sperm bristle category", fill = "Inferred mating syndrome") +
    #theme(legend.position = "none")
    theme(legend.position = c(0.05, 0.8), legend.direction="horizontal") +
    guides(fill=guide_legend(override.aes=list(shape=21)))
  legend_pPCA_inferred <- get_legend(pPCA_inferred)
  pPCA_inferred
  
  ## PCA Result with inferred syndrome and clade combined
  
  pPCA_inferred1 <- ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
    geom_point(aes(shape= factor(ppca1[[1]]$clade), fill=factor(ppca1[[1]]$inferred_syndrome)), 
               size=size, stroke = stroke, alpha=alpha) +
    scale_fill_manual(labels = c("reciprocal", "intermediate", "hypodermic", "no observation"), values = palette2) + 
    scale_shape_manual(labels = c("absent", "reduced", "present"), values = c(24,21)) + 
    labs(shape = "sperm bristle category", fill = "Inferred mating syndrome") +
    #theme(legend.position = "none")
    theme(legend.position = c(0.05, 0.8), legend.direction="horizontal") +
    guides(fill=guide_legend(override.aes=list(shape=21)))
  legend_pPCA_inferred1 <- get_legend(pPCA_inferred1)
  
  pPCA_inferred1 + theme(legend.position = "none")
  
  #################
  
  
  plot_grid(pPCA_clade, pPCA_p1, pPCA_inferred,ncol=1)
  plot_grid( pPCA_p1, pPCA_inferred + geom_text_repel(aes(label = ppca1[[1]]$Short),
              nudge_x = 0.05, nudge_y = 0.02, size = 3),ncol=1)
  
  heatp <-   pPCA_1_L %>%
  gather(key = PC, value = val, PC1:PC5) %>%
    mutate(var = factor(var, levels = rev(var_order))) %>%
  ggplot(aes(x = PC, y = var)) +
  geom_tile(aes(fill=val)) +
  geom_text(aes(label=round(val,2)), size =9) + 
  scale_fill_gradient2(low="royalblue", mid="white", high="firebrick1") +
    theme(legend.position = "none", 
          axis.title = element_blank())
  print(heatp)
  
  
  heatp_small <-   pPCA_1_L %>%
  gather(key = PC, value = val, PC1:PC2) %>%
    mutate(var = factor(var, levels = rev(var_order))) %>%
  ggplot(aes(x = PC, y = var)) +
  geom_tile(aes(fill=val)) +
  geom_text(aes(label=round(val,2)), size =9) + 
  scale_fill_gradient2(low="royalblue", mid="white", high="firebrick1") +
  theme(legend.position = "none", 
          axis.title = element_blank())
  print(heatp_small)
  
  
  var_exp <- diag(pPCA_1$Eval)/sum(pPCA_1$Eval)*100
  df <- data.frame(var_explained=var_exp,
                   cum_var_explained = cumsum(var_exp))
  df$PC <- rownames(df)
  df$PC <- factor(df$PC, levels = df$PC)
  pl <- ggplot(df) +
  geom_bar(aes(x=PC, var_explained), stat = "identity") +
  geom_text(aes(x=PC, y= -0.5, label=round(cum_var_explained,1)))
  print(pl)
  
  ##
  closeAllConnections()
  pdf(paste0("fig/pPCA_selected_traits_", names(trees)[[i]],".pdf"),width = 10, height = 10, useDingbats = FALSE)
  print(pPCA_plane + theme(legend.position = "none"))
  print(pPCA_clade + theme(legend.position = "none"))
  print(pPCA_p1 + theme(legend.position = "none"))
  print(pPCA_p1 + theme(legend.position = "none") + geom_text_repel(aes(label = ppca1[[1]]$Short),
              nudge_x = 0.05, nudge_y = 0.02, size = 3))
  print(pPCA_inferred + theme(legend.position = "none"))
  print(pPCA_inferred1 + theme(legend.position = "none"))
  print(plot_grid(
     legend_pPCA_plane,
     legend_pPCA_clade,  
     legend_pPCA_p1,
     legend_pPCA_inferred
    ))
  print(heatp)
  print(heatp_small)
  print(pl)
  dev.off()
  
  pdf(paste0("fig/pPCA_selected_traits_inferred_", names(trees)[[i]],".pdf"),width = 10, height = 7, useDingbats = FALSE)
  print(pPCA_inferred + theme(legend.position = "none"))
  print(pPCA_inferred1 + theme(legend.position = "none"))
  dev.off()
  
  ####################
  ggplot(as.data.frame(pPCA_1$S), aes(x = PC1, y=PC2)) +
    geom_point(aes(color= factor(ppca1[[1]]$clade)), size=8, shape = 1) +
    geom_point(aes(color= factor(ppca1[[1]]$bristle), shape=factor(ppca1[[1]]$Rec)), size=4)  
  
  pPCA_p1
  pPCA_p1_lab <- pPCA_p1 + geom_text_repel(aes(label = ppca1[[1]]$Short),
              nudge_x = 0.05, nudge_y = 0.02, size = 3)
  
  pdf(paste0("fig/PCA_no_cilia_removeNA_", names(trees)[[i]],".pdf"), width=14, height=10)
  print(pPCA_p1)
  print(pPCA_p1 + 
    geom_text(data=pPCA_1_L, aes(x=PC1, y=PC2,  label = rownames(pPCA_1$L))) +
    geom_segment(data=pPCA_1_L, aes(x=0, y=0, xend=PC1, yend=PC2),
                 arrow=arrow(length = unit(0.01, "npc")),
                 size=1, color="grey", alpha = 0.8))
  
  print(pPCA_p1_lab + 
    geom_text(data=pPCA_1_L, aes(x=PC1, y=PC2,  label = rownames(pPCA_1$L))) +
    geom_segment(data=pPCA_1_L, aes(x=0, y=0, xend=PC1, yend=PC2),
                 arrow=arrow(length = unit(0.01, "npc")),
                 size=1, color="grey", alpha = 0.8))
  
  pPCA_p1_e <- pPCA_p1 + 
  stat_ellipse(aes(color=factor(ppca1[[1]]$bristle),
                   group=factor(ppca1[[1]]$bristle)),
                   level = 0.95, geom = "polygon",
               alpha = 0, type = "t", size=2) +
  scale_color_manual(labels = c("absent", "reduced", "present"), values = cols) 

    legend1 <- get_legend(pPCA_p1_e)
    desc <- "95% confidence ellipse assuming t-distribution"
    pf1 <- pPCA_p1_e +theme(legend.position="none")
  pPCA_p1_format <- pf1 + draw_label(desc, 0.6, 1.45) + ggtitle("PCA categorical data - removed species with missing data")
  print(pPCA_p1_format)
  
  var_exp <- diag(pPCA_1$Eval)/sum(pPCA_1$Eval)*100
  df <- data.frame(var_explained=var_exp,
                   cum_var_explained = cumsum(var_exp))
  df$PC <- rownames(df)
  df$PC <- factor(df$PC, levels = df$PC)
  pl <- ggplot(df) +
  geom_bar(aes(x=PC, var_explained), stat = "identity") +
  geom_text(aes(x=PC, y= -0.5, label=round(cum_var_explained,1)))
  print(pl)
  dev.off()
  
  # output data
  lambda <- round(pPCA_1$lambda,2)
  pPCA_1_S$lambdaPCA <- lambda
  write.csv(pPCA_1_S,
            paste0("output/PCA_no_cilia_removeNA_values_",
                   names(trees)[[i]], ".csv"),
            row.names = FALSE)
  write.csv(pPCA_1_L, paste0("output/PCA_no_cilia_removeNA_loadings_",
                             names(trees)[[i]], ".csv"),
            row.names = FALSE)
  }
```

phylomorphospace plot for main figure

```{r}
ignore = c("Rec", "bristle", "clade", "inferred_syndrome")
  
ppca1 <- main_df %>%
  select(Short, Rec, bristle, clade, inferred_syndrome,
         # stylet traits
         sty_avg_um_log_mean,
         wt_um_log_mean,
         wb_um_log_mean,
         sty_curveSum_avg_mean,
         #sty_curveSum_scaled_mean,
         delta_e_um_log_mean,
         sharp,
         # sperm traits
         total_sprmLen_um_log_mean,
         brstlLen_um_per_sperm_log_mean,
         velum,
         brush,
         # antrum traits
         ant_thickness,
         ant_cellular_valve,
         ant_chambers,
         No_openings,
         # body size
         bdSize_um_logsqrt_mean) %>%
  mutate(
    brush = ifelse(brush %in% c(1,2),1, ifelse(!is.na(brush),0,NA))
    ) %>% 
  rename(
    'stylet length (logum)' = sty_avg_um_log_mean,
    'width distal stylet opening (logum)' = wt_um_log_mean,
    'width proximal stylet opening (logum)' = wb_um_log_mean,
    'curviness of stylet' = sty_curveSum_avg_mean,
    #'curviness of stylet' = sty_curveSum_scaled_mean,
    'asymetry of distal thickening (logum)' = delta_e_um_log_mean,
    'sharpness of stylet' = sharp,
    # sperm traits
    'sperm length (logum)'=total_sprmLen_um_log_mean,
    'sperm bristle length (logum)' = brstlLen_um_per_sperm_log_mean,
    'sperm velum' = velum,
    'sperm brush' = brush,
    # antrum traits
    'antrum thickness' = ant_thickness,
    'antrum cellular valve' = ant_cellular_valve,
    'complexity of antrum chamber' = ant_chambers,
    'number of vaginas' = No_openings,
         # body size
    'body area (logsqrtum)' = bdSize_um_logsqrt_mean
    )%>%
  as.data.frame() %>%
prepare_phyl_pca(tree = trees[[i]], ignore=ignore)

var_order <- c('stylet length (logum)',
    'curviness of stylet',
    'width proximal stylet opening (logum)',
    'width distal stylet opening (logum)',
    'asymetry of distal thickening (logum)',
    'sharpness of stylet',
    'sperm length (logum)',
    'sperm bristle length (logum)',
    'sperm velum',
    'sperm brush',
    'antrum thickness',
    'antrum cellular valve',
    'complexity of antrum chamber',
    'number of vaginas',
    'body area (logsqrtum)')
# remove the ignored columns which are just used to paint later
tmp_df <- ppca1[[1]][,setdiff(names(ppca1[[1]]), c(treename, ignore))]
t <- ppca1[[2]]

cols <- c("black", "#1FBDCC", "#FF0E98", "grey")

pPCA_1 <- phyl.pca(t, tmp_df, method = "lambda", mode="corr")

  # I want the majority of the loadings to be positive
if (sum(-1*pPCA_1$L[,1]) > sum(pPCA_1$L[,1])){
  pPCA_1$L[,1] <- -1*pPCA_1$L[,1]
  pPCA_1$S[,1] <- -1*pPCA_1$S[,1]
}

if (sum(-1*pPCA_1$L[,2]) > sum(pPCA_1$L[,2])){
  pPCA_1$L[,2] <- -1*pPCA_1$L[,2]
  pPCA_1$S[,2] <- -1*pPCA_1$S[,2]
}

pPCA_1_L <- as.data.frame(pPCA_1$L) %>%
mutate(var = rownames(pPCA_1$L))

pPCA_1_S <- as.data.frame(pPCA_1$S)  %>%
mutate(Short = rownames(pPCA_1$S))

ppca1[[1]]$Rec <- ifelse(is.na(ppca1[[1]]$Rec), "no_observation", ppca1[[1]]$Rec)

## morphospace plot
# color by label
ppca1[[1]]$inferred_syndrome[match(ppca1[[2]]$tip.label, ppca1[[1]]$Short)]
colmap <- setNames(palette2[1:length(unique(ppca1[[1]]$inferred_syndrome))],
                   sort(unique(ppca1[[1]]$inferred_syndrome)))
ordered_trait <- ppca1[[1]]$inferred_syndrome[match(ppca1[[2]]$tip.label,
                                                    ppca1[[1]]$Short)]

tipcols <- colmap[match(ordered_trait, names(colmap))]
names(tipcols) <- ppca1[[2]]$tip.label
t_cols<-c(tipcols[ppca1[[2]]$tip.label],rep("black",ppca1[[2]]$Nnode))
names(t_cols)<-1:(length(ppca1[[2]]$tip)+ppca1[[2]]$Nnode)

tree=ppca1[[2]]
X=pPCA_1$S[,1:2]
nsteps = 200
direction="to"
sleep=0
lab = "off"
#tree <- minRotate(reorder(tree, "cladewise"), X[, 2], print = FALSE)
X <- X[tree$tip.label, ]
A <- cbind(fastAnc(tree, X[, 1]), fastAnc(tree, X[, 2]))
cladogram <- tree
cladogram$edge.length <- NULL
mar <- par()$mar
plotTree(cladogram, type = "cladogram", nodes = "centered", 
         plot = FALSE)
obj <- get("last_plot.phylo", envir = .PlotPhyloEnv)
obj$xx <- obj$xx * (max(c(X[, 1], A[, 1])) - min(c(X[, 1], 
                                                   A[, 1]))) + min(c(X[, 1], A[, 1]))
xlim <- range(obj$xx)
xlim[2] <- xlim[2] + abs(diff(xlim)) * (obj$x.lim[2] - 1)
obj$yy <- (obj$yy - min(obj$yy))/(max(obj$yy) - min(obj$yy)) * 
  (max(c(X[, 2], A[, 2])) - min(c(X[, 2], A[, 2]))) + min(c(X[, 
                                                              2], A[, 2]))
ylim <- range(obj$yy)
X0 <- cbind(obj$xx[1:Ntip(tree)], obj$yy[1:Ntip(tree)])
rownames(X0) <- tree$tip.label
A0 <- cbind(obj$xx[1:tree$Nnode + Ntip(tree)], obj$yy[1:tree$Nnode + 
                                                        Ntip(tree)])
rownames(A0) <- 1:tree$Nnode + Ntip(tree)


edge_length <- length(tree$edge[,2])
xx <- edge_length-179
e_cols <- c(rep("black", edge_length-xx), rep("grey", xx))
names(e_cols) <- tree$edge[,2]
if (direction == "both") 
  nsteps <- ceiling(nsteps/2)
xlim=c(-2,6)
for (i in 0:nsteps) {

  if (i%%100 == 0) {
  pdf(paste0("fig/phymorph/",i,".pdf"), width = 14, height = 14)
  phylomorphospace(tree, ((nsteps - i) * X0 + i * X)/nsteps, 
                   ((nsteps - i) * A0 + i * A)/nsteps, xlim = xlim, 
                   ylim = ylim, label = lab,
                   control=list(col.node=t_cols, col.edge=e_cols
                                    ), node.size=c(0.0,4),
                       xlab="PC1",ylab="PC2",lwd=3) # add ... support again
  dev.off()
  }
}

```


```{r}
project.phylomorphospace_mod(ppca1[[2]], pPCA_1$S[,1:2],
                           direction = "to", sleep=0,
                           control=list(col.node=t_cols
                                        ), node.size=c(0.8,1.8),
                           xlab="PC1",ylab="PC2")
```
