---
title: "18_inferred_vs_sperm_length"
author: "Jeremias Brand"
date: "12 Januar 2020"
output: html_document
---


```{r}
require(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = "..")
```


```{r results="hide", warning=FALSE }

source("R/functions.R")
library(dplyr)
library(scales)
library(tidyr)
library(picante)
library(ggplot2)
library(readxl)
library(writexl)
library(purrr)
library(ape)
library(phytools)
library(geiger)
library(phangorn)
library(forcats)
library(viridis)
library(ggtree)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(gtable)
library(ggstance)
library(stringr)
library(ICC)
library(lmtest)
library(nlme)
library(bayou)
library(ggrepel)
library(plotly)
library(cowplot)
library(rr2)

cols <- c("black", "lightblue", "blue", "Grey")
shapes <- c(19,17,15,1)
treename = "Short"


pr = "#006837"
int = "#78c679"
ab = "#ffffcc"

palette1 <- c(pr, int, ab, "grey")
palette2 <- c(ab, int, pr, "grey")
palette3 <- c(ab, pr, "grey")
# Prevent MASS and dplyr clashes
select <- dplyr::select
```

## Load data

```{r}
## Load data

m_df <- readxl::read_excel("Tab_S2.xlsx", sheet = 2, na = c("","NA"),
                                 col_types = c("text", "text", rep("numeric",8)))


s_df <- readxl::read_excel("Tab_S3.xlsx", sheet = 2, na = c("","NA"),
                                  col_types = c(
                                    rep("text",5),
                                    rep("numeric",10),
                                    "text",
                                    rep("numeric",12),
                                    "text"
                                    )
                                 )

N_df <- readxl::read_excel("Tab_S1.xlsx", sheet = 2) %>%
  mutate(
    Species = gsub("M\\.", "Macrostomum", Species)
  )
```

## Transform the data

We want to do mean(log10(x)) *NOT* log(mean(x)). 
Since the former is a much better representation of a species value.


```{r}
morpho_df <- m_df %>%
  group_by(sci_name, ID) %>%
  mutate_at(c("body_size"), funs(body_size_logsqrt = log10(sqrt(.)))) %>%
  mutate_at(c("stylet_length", "stylet_proximal", "stylet_distal", "sperm_length", "bristle_length"),
            funs(log = log10(.))) %>%
  mutate_at(c("stylet_curvyness", "bristle_length"), funs(log = log10(. + 1)))

morpho_df2 <- morpho_df %>%
  group_by(sci_name) %>%
  select(sci_name, contains("log")) %>%
  summarise_if(is.numeric, 
               funs(mean = mean(., na.rm = TRUE), 
                       N = sum(!is.na(.)))
               ) %>%
  mutate( sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", sci_name)))
```

## LOAD TREES

```{r}
trees <- readRDS(file = paste0("trees.Rdata"))
ltre <- length(trees)

# Color the hypodermic clade
t_injecting_clade <- extract.clade(trees[[3]],154)
names_injecting <- t_injecting_clade$tip.label

s_df$clade <- ifelse(s_df$Short %in% names_injecting, "clade1", "clade2")

species_df <- s_df %>%
  mutate(sci_name_short = gsub(" ", "_", gsub("Macrostomum", "M.", Species)),
         inferred_syndrome_2states = case_when(
           inferred_syndrome == "intermediate" ~ "hypodermic",
           TRUE ~ inferred_syndrome)
         ) %>%
  select(-contains("mean"))

# Some in the species list have not been measured
species_df$Species %in% morpho_df2$sci_name
sum(species_df$Species %in% morpho_df2$sci_name)

# All measured are in the species list
morpho_df2$sci_name %in% species_df$Species

# Combine them
main_df <- left_join(species_df, morpho_df2, by = "sci_name_short")

length(species_df$Short)
species_df$Short %in% trees[[3]]$tip.label

```

## Color palette

```{r}
Nclusters=3
cl <- parallel::makeCluster(Nclusters)
doParallel::registerDoParallel(cl)
library(foreach)

pr = "#006837"
int = "#78c679"
ab = "#ffffcc"

palette1 <- c(pr, int, ab, "grey")
palette2 <- c(ab, int, pr, "grey")
palette3 <- c(ab, pr, "grey")
```


### multiply trees by 100 

The ML optimizer can run into issues due to numeric underflow. 

```{r}
trees <- lapply(trees,function(tre){tre$edge.length <- tre$edge.length * 100; return(tre)})
```

Make sure we are using factors

```{r}
main_df <- main_df %>%
  mutate(Rec_binary = factor(Rec_binary, levels = c(0,1)),
         bristle_binary = factor(bristle_binary, levels = c(0,1)),
         inferred_syndrome_2states = factor(inferred_syndrome_2states, 
                                            levels = c("hypodermic","reciprocal")))
```

## Effect of sperm length

Model comparison should be done using ML not REML.
Model fitting however should be done using REML because ML is biased especially if
number of parameters (p) is similar in size to number of observations (N).

### Setup formulas and switches

```{r}
outname <- "EL_sperm_length_cat"
FILTER = FALSE
DIAGNOSTIC = FALSE
#syndrome = "hypodermic"

pred_name <- list("received sperm binary",
                  "bristle state binary",
                  "antrum binary",
                  "inferred syndrome binary")
preds <- list("Rec_binary",
              "bristle_binary",
              "an_binary",
              "inferred_syndrome_2states")
formulas <- list("sperm_length_log_mean ~ Rec_binary",
                 "sperm_length_log_mean ~ bristle_binary",
                 "sperm_length_log_mean ~ an_binary",
                 "sperm_length_log_mean ~ inferred_syndrome_2states")
N = "sperm_length_log_N"
N_var = "~1/sperm_length_log_N"
dependent_var = "sperm_length_log_mean"
cor_form = "~ Short"
```

### Setup ouput lists

```{r}
main_res  <-  list()
main_res_fits <- list()
main_res_models <- list()
```

### Model comparison using ML

```{r}
for (i in 1:length(formulas)){
  for (j in 1:ltre){
    predictor <- pred_name[[i]]
    df <- main_df
    if(FILTER == TRUE){
      df <- filter(main_df, inferred_syndrome_2states == syndrome)
    }
    comp_df <- df %>%
      select(Short, preds[[i]], dependent_var, N) %>%
      as.data.frame() %>%
      prepare_phyl_pca(tree = trees[[j]], treename = "Short")
    
    BM <- gls(as.formula(formulas[[i]]),
              data = comp_df[[1]], 
              correlation = corBrownian(1,
                                        phy=comp_df[[2]],
                                        form = as.formula(cor_form)), 
              weight = as.formula(N_var), method = "ML")
    
    lambda <- tryCatch(
      {gls(as.formula(formulas[[i]]),
           data = comp_df[[1]],
           correlation = corPagel(1,
                                  phy=comp_df[[2]],
                                  form = as.formula(cor_form)),
           weight = as.formula(N_var), method = "ML",
           control=list(maxIter=500))},
      error=function(cond){
        
        print(cond)
        
        optimal_lambda <- optimize_lambda(
          formula=as.formula(formulas[[i]]),
          data = comp_df[[1]],
          phy=comp_df[[2]],
          weight = as.formula(N_var),
          method = "ML",
          form = as.formula(cor_form),
          control=list())
        
        gls(as.formula(formulas[[i]]),
            data = comp_df[[1]], 
            correlation = corPagel(optimal_lambda, 
                                   phy=comp_df[[2]],
                                   fixed=TRUE,
                                   form = as.formula(cor_form)), 
            weight = as.formula(N_var),
            method = "ML",
            control=list(maxIter=500))
      }
    )
    
    # limit lambda to 0 - 1
    if(lambda$modelStruct[[1]] < 0){
      lambda <- gls(as.formula(formulas[[i]]),
                    data = comp_df[[1]], 
                    correlation = corPagel(0, 
                                           phy=comp_df[[2]],
                                           form = as.formula(cor_form),
                                           fixed=TRUE), 
                    weight = as.formula(N_var), method = "ML",
                    control=list(maxIter=500))
    } else if (lambda$modelStruct[[1]] > 1){
      # this means it is a BM model
      lambda <- gls(as.formula(formulas[[i]]),
                    data = comp_df[[1]], 
                    correlation = corPagel(1, 
                                           phy=comp_df[[2]], 
                                           form = as.formula(cor_form), 
                                           fixed=TRUE), 
                    weight = as.formula(N_var),
                    method = "ML",
                    control=list(maxIter=500)) 
    }
    # convergence issues
    OU <- tryCatch(
      {gls(as.formula(formulas[[i]]),
           data = comp_df[[1]],
           correlation = corMartins(1,
                                    phy=comp_df[[2]],
                                    form = as.formula(cor_form)),
           weight = as.formula(N_var),
           method = "ML",
           control=list(maxIter=500))},
      error=function(cond){
        
        print(cond)
        optimal_alpha <- optimize_alpha(
          formula = as.formula(formulas[[i]]),
          data = comp_df[[1]],
          phy=comp_df[[2]],
          weight = as.formula(N_var),
          method = "ML",
          form = as.formula(cor_form),
          control=list())
        
        gls(as.formula(formulas[[i]]),
            data = comp_df[[1]], 
            correlation = corMartins(optimal_alpha, 
                                     phy=comp_df[[2]],
                                     form = as.formula(cor_form),
                                     fixed=TRUE), 
            weight = as.formula(N_var),
            method = "ML",
            control = list(maxIter=500))
      })
    ############
    # Summarise fits and calculate AICc
    models <- list(BM=BM,lambda=lambda,OU=OU)
    fits <- data.frame(rbind(AICc_pgls_nlme(BM),
                             AICc_pgls_nlme(lambda), AICc_pgls_nlme(OU)))
    fits$predictor <- predictor
    fits$model <-  c("BM", "lambda", "OU")
    fits$tree <- names(trees)[[j]]
    fits$N <- BM$dims$N
    fits$modelPar <- round(
      as.numeric(c(NA, lambda$modelStruct[[1]], OU$modelStruct[[1]])),2)
    # Models where optimization failed will be ignored
    fits$AIC <- ifelse(fits$AIC == -Inf, NA, fits$AIC)
    fits$AICc <- ifelse(fits$AICc == -Inf, NA, fits$AICc)
    # look at the results of the best fitting model
    best_fit <- na.omit(fits$model[fits$AICc == min(fits$AICc, na.rm = TRUE)])
    # if lambda is 1 then BM and lambda will be the same model
    # we can then choose lambda to display
    if(length(best_fit) > 1 & sum(best_fit == c("BM", "lambda")) == 2) {
      best_fit = "lambda"
    }
    out_fits <- fits %>%
      select(predictor, tree, N, model, modelPar, AIC, AICc)
    ############
    main_res_fits <- append(main_res_fits,list(out_fits))
    
  }
}
```

### Output ML Fits

```{r}

write.csv(do.call(rbind, main_res_fits),
          file=paste0("output/PGLS_",outname,"_weighted_fit.csv"))
```

### Fit models using REML

```{r}
for (i in 1:length(formulas)){
for (j in 1:ltre){
  predictor <- pred_name[[i]]
  df <- main_df
  if(FILTER == TRUE){
    df <- filter(main_df, inferred_syndrome_2states == syndrome)
  }
  comp_df <- df %>%
    select(Short, preds[[i]],
           dependent_var, N
           ) %>%
  as.data.frame() %>%
  prepare_phyl_pca(tree = trees[[j]], treename = "Short")
  
  
  lambda <- tryCatch(
    {gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(1, phy=comp_df[[2]], form = ~Short), 
                 weight = as.formula(N_var), method = "REML",
                control=list(maxIter=500))},
    error=function(cond){
      print(cond)
      optimal_lambda <- optimize_lambda(
        formula=as.formula(formulas[[i]]), data = comp_df[[1]], phy=comp_df[[2]],
        weight = as.formula(N_var), method = "REML", control=list())
      gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(optimal_lambda, 
                                        phy=comp_df[[2]], form = ~Short, fixed=TRUE), 
                 weight = as.formula(N_var), method = "REML",
                 control=list(maxIter=500))})
  
  # limit lambda to 0 - 1
  if(lambda$modelStruct[[1]] < 0){
    lambda <- gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(0, phy=comp_df[[2]], form = ~Short, fixed=TRUE), 
                 weight = as.formula(N_var), method = "REML",
                 control=list(maxIter=500))
  } else if (lambda$modelStruct[[1]] > 1){
    # this means it is a BM model
     lambda <- gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(1, phy=comp_df[[2]], form = ~Short, fixed=TRUE), 
                 weight = as.formula(N_var), method = "REML",
                 control=list(maxIter=500)) 
  }

  corStructure = lambda$modelStruct[[1]]
  mod0 <- gls(as.formula("sperm_length_log_mean ~ 1"), data = comp_df[[1]], 
                 correlation = corPagel(1,
                                        phy=comp_df[[2]], form = ~Short),
                 weight = as.formula(N_var), method = "REML",
                control=list(maxIter=500))
  
  full =   lambda 
  phy_only = mod0
  data_only = lm(as.formula(formulas[[i]]), data = comp_df[[1]])
  R2_pred_data = round(R2.pred(full, phy_only),3)
  R2_pred_phy = round(R2.pred(full, data_only),3)
  R2_pred_full = round(R2.pred(full),3)
  
  #####################
  s_best <- summary(lambda)
  res_df <- data.frame(s_best$tTable)
  res_df$Estimate <- rownames(s_best$tTable)
  res_df$model <- "lambda"
  res <- res_df
  res$R2pred <- c(R2_pred_full, rep(R2_pred_data, nrow(res)-1))
  
  res$predictor = pred_name[[i]]
  res$phylogeny = names(trees)[[j]]
  res$lambda = lambda$modelStruct[[1]]
  res$N = nrow(comp_df[[1]])
  #############
  # DIAGNOSTICS
  if (DIAGNOSTIC == TRUE) {
    png(paste0("fig/diag/",outname, "_", names(trees)[j],"_",preds[i],".png"),
      width = 7, height = 7, unit="in", res = 300)
    pgls_diag_gls(name=paste0(names(trees)[j]," ",predictor), mod_type = "lambda",
                model=lambda, formula=as.formula(formulas[[i]]),
                data = comp_df[[1]], phy=comp_df[[2]], 
                weight = as.formula(N_var), method = "REML", control=list(maxIter=500, opt="optim"))
    dev.off()
  }

  ############
  main_res <- append(main_res,list(res))
  main_res_models <- append(main_res_models, list(lambda))
}
}

main_res <- do.call(rbind, main_res)
main_res$p.value <- format.pval(main_res$p.value, eps=0.001, digits=2)
main_res_out <- main_res %>%
  select(predictor, phylogeny, Estimate, N, lambda, Value, Std.Error, t.value, p.value, R2pred) %>%
  mutate_if(is.numeric, round, 2)
```

### Output Table and Models

```{r}
write.csv(main_res_out, file=paste0("output/PGLS_",outname,"_weighted_PGLS.csv"))

names(main_res_models) <- apply(expand.grid(names(trees), preds), 1, paste, collapse="_")
saveRDS(main_res_models, paste0("output/PGLS_", outname,"_weighted_PGLS.RDS"))
```


## Plotting sperm length 

This is a combination of the plot in the main Figure and the Supplement.

```{r}
list("sperm_length_log_mean ~ Rec_binary",
     "sperm_length_log_mean ~ bristle_binary",
     "sperm_length_log_mean ~ an_binary",
     "sperm_length_log_mean ~ inferred_syndrome_2states")
s <- c(seq(30,60,10),80,100,120,140)
w <- 0.01
ann_size = 4
f = 2.4
t = 2.2

plot_df <- main_df


rows <-  main_res_out$predictor == "received sperm binary" &
  main_res_out$phylogeny == "C-IQ-TREE" 

r_df <- main_res_out[rows,]

lab <- c(
  bquote(N==.(r_df[1,4])),
  bquote(lambda==.(r_df[1,5])),
  bquote(beta==.(r_df[2,6])),
  bquote(t==.(r_df[2,8])),
  bquote(p~.(r_df[2,9])),
  bquote(R[pred]^2==.(r_df[1,10]))
)

p_Rec_binary <- ggplot(main_df[main_df$Rec_binary %in% c(0,1), ],
                       aes(x=Rec_binary, y = sperm_length_log_mean)) + 
  geom_boxplot(aes(fill= factor(Rec_binary))) +
  geom_jitter(height=0, width=0.01, size=3, color="grey30") +
  scale_fill_manual(values=palette3) + 
  scale_x_discrete(labels=c("hypodermic", "antrum only")) +
  scale_y_continuous(breaks = log10(s), label=s) +
  xlab("Received sperm location") +
  ylab(expression("Sperm length"~(""*mu*"m"))) +
  theme_cowplot() +
  theme(legend.position = "none") + 
  annotate("text",
           x = 0.6,
           y = seq(from = f, to = t, length.out = length(lab)),
           label = lab ,
           parse = TRUE,
           size = ann_size,
           hjust = 0)


p_Rec_binary 


rows <-  main_res_out$predictor == "bristle state binary" &
  main_res_out$phylogeny == "C-IQ-TREE" 

r_df <- main_res_out[rows,]

lab <- c(
  bquote(N==.(r_df[1,4])),
  bquote(lambda==.(r_df[1,5])),
  bquote(beta==.(r_df[2,6])),
  bquote(t==.(r_df[2,8])),
  bquote(p~.(r_df[2,9])),
  bquote(R[pred]^2==.(r_df[1,10]))
)

p_bristle_binary <- ggplot(main_df[main_df$bristle_binary %in% c(0,1), ],
                       aes(x=factor(bristle_binary), y = sperm_length_log_mean)) + 
  geom_boxplot(aes(fill= factor(bristle_binary))) +
  geom_jitter(height=0, width=w, size=3, color="grey30") +
  scale_fill_manual(values=palette3) + 
  scale_x_discrete(labels=c("absent/reduced", "present")) +
  scale_y_continuous(breaks = log10(s), label=s) +
  xlab("Bristle state") +
  ylab(expression("Sperm length"~(""*mu*"m"))) +
  theme_cowplot() +
  theme(legend.position = "none") + 
  annotate("text",
           x = 0.6,
           y = seq(from = f, to = t, length.out = length(lab)),
           label = lab ,
           parse = TRUE,
           size = ann_size,
           hjust = 0)


# antrum

rows <-  main_res_out$predictor == "antrum binary" &
  main_res_out$phylogeny == "C-IQ-TREE" 

r_df <- main_res_out[rows,]

lab <- c(
  bquote(N==.(r_df[1,4])),
  bquote(lambda==.(r_df[1,5])),
  bquote(beta==.(r_df[2,6])),
  bquote(t==.(r_df[2,8])),
  bquote(p~.(r_df[2,9])),
  bquote(R[pred]^2==.(r_df[1,10]))
)

p_an_binary <- main_df[main_df$an_binary %in% c(0, 1), ] %>% 
  mutate(an_binary = factor(an_binary, levels=c(0, 1))) %>% 
           ggplot(aes(x = an_binary, y = sperm_length_log_mean)) + 
  geom_boxplot(aes(fill=an_binary), outlier.shape = NA) +
  geom_jitter(height=0, width=0.01, size=3, color="grey30") +
  scale_fill_manual(values=palette3) + 
  scale_x_discrete(labels=c("thin", "thickened")) +
  scale_y_continuous(breaks = log10(s), label=s) +
  xlab("Antrum type") +
  ylab(expression("Sperm length"~(""*mu*"m"))) +
  theme_cowplot() +
  theme(legend.position = "none") + 
  annotate("text",
           x = 0.6,
           y = seq(from = f, to = t, length.out = length(lab)),
           label = lab ,
           parse = TRUE,
           size = ann_size,
           hjust = 0)


# syndrome
rows <-  main_res_out$predictor == "inferred syndrome binary" &
  main_res_out$phylogeny == "C-IQ-TREE" 

r_df <- main_res_out[rows,]

lab <- c(
  bquote(N==.(r_df[1,4])),
  bquote(lambda==.(r_df[1,5])),
  bquote(beta==.(r_df[2,6])),
  bquote(t==.(r_df[2,8])),
  bquote(p~.(r_df[2,9])),
  bquote(R[pred]^2==.(r_df[1,10]))
)


p_inferred_syndrome_binary <- main_df[main_df$inferred_syndrome_2states %in% c("hypodermic","reciprocal"),] %>%
  ggplot(aes(x=inferred_syndrome_2states,y = sperm_length_log_mean)) + 
  geom_boxplot(aes(fill= inferred_syndrome_2states), outlier.shape = NA) +
  geom_jitter(height=0, width=w, size=3, color="grey30") +
  scale_fill_manual(values=palette3) + 
  scale_y_continuous(breaks = log10(s), label=s) +
  xlab("Inferred syndrome") +
  ylab(expression("Sperm length"~(""*mu*"m"))) +
  theme_cowplot() +
  theme(legend.position = "none") + 
  annotate("text",
           x = 0.6,
           y = seq(from = f, to = t, length.out = length(lab)),
           label = lab ,
           parse = TRUE,
           size = ann_size,
           hjust = 0)

p_inferred_syndrome_binary_main <- main_df[main_df$inferred_syndrome_2states %in% c("hypodermic","reciprocal"),] %>%
  ggplot(aes(x=inferred_syndrome_2states,y = sperm_length_log_mean)) + 
  geom_boxplot(aes(fill= inferred_syndrome_2states), outlier.shape = NA, width = 0.4) +
  geom_jitter(height=0, width=w, size=3, color="grey30") +
  scale_fill_manual(values=palette3) + 
  scale_y_continuous(breaks = log10(s), label=s) +
  xlab("Inferred syndrome") +
  ylab(expression("Sperm length"~(""*mu*"m"))) +
  theme_cowplot() +
  theme(legend.position = "none") + 
  annotate("text",
           x = 0.6,
           y = seq(from = f, to = t, length.out = length(lab)),
           label = lab ,
           parse = TRUE,
           size = ann_size,
           hjust = 0)





pdf("fig/sperm_length_boxplot_PGLS.pdf", width=14, useDingbats = FALSE)
plot_grid(p_Rec_binary,p_bristle_binary, p_an_binary, p_inferred_syndrome_binary, nrow=1,
          labels = c("A", "B", "C", "D"), label_size = 20)
dev.off()

pdf("fig/sperm_length_boxplot_PGLS_supplement.pdf", width=11, useDingbats = FALSE)
plot_grid(p_Rec_binary,p_bristle_binary, p_an_binary, nrow=1,
          labels = c("A", "B", "C"), label_size = 20)
dev.off()


pdf("fig/sperm_length_boxplot_PGLS_main.pdf", width=3, useDingbats = FALSE)
plot_grid(p_inferred_syndrome_binary_main, nrow=1)
dev.off()

```


## Including body size as a covariate

### Setup formulas and switches

```{r}
outname <- "EL_sperm_length_body_size"
FILTER = FALSE
DIAGNOSTIC = FALSE
#syndrome = "hypodermic"

pred_name <- list("received sperm binary",
                  "bristle state binary",
                  "antrum binary",
                  "inferred syndrome binary")
preds <- list("Rec_binary",
              "bristle_binary",
              "an_binary",
              "inferred_syndrome_2states")
formulas <- list("sperm_length_log_mean ~ Rec_binary + body_size_logsqrt_mean",
                 "sperm_length_log_mean ~ bristle_binary + body_size_logsqrt_mean",
                 "sperm_length_log_mean ~ an_binary + body_size_logsqrt_mean",
                 "sperm_length_log_mean ~ inferred_syndrome_2states + body_size_logsqrt_mean")
N = "sperm_length_log_N"
N_var = "~1/sperm_length_log_N"
dependent_var = "sperm_length_log_mean"
cor_form = "~ Short"
```

### Setup ouput lists

```{r}
body_res  <-  list()
body_res_fits <- list()
body_res_models <- list()
```


```{r}
### Fit models using REML

for (i in 1:length(formulas)){
for (j in 1:ltre){
  predictor <- pred_name[[i]]
  df <- main_df
  if(FILTER == TRUE){
    df <- filter(main_df, inferred_syndrome_2states == syndrome)
  }
  comp_df <- df %>%
    select(Short, preds[[i]],
           dependent_var, N, body_size_logsqrt_mean
           ) %>%
  as.data.frame() %>%
  prepare_phyl_pca(tree = trees[[j]], treename = "Short")
  
  
  lambda <- tryCatch(
    {gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(1, phy=comp_df[[2]], form = ~Short), 
                 weight = as.formula(N_var), method = "REML",
                control=list(maxIter=500))},
    error=function(cond){
      print(cond)
      optimal_lambda <- optimize_lambda(
        formula=as.formula(formulas[[i]]), data = comp_df[[1]], phy=comp_df[[2]],
        weight = as.formula(N_var), method = "REML", control=list())
      gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(optimal_lambda, 
                                        phy=comp_df[[2]], form = ~Short, fixed=TRUE), 
                 weight = as.formula(N_var), method = "REML",
                 control=list(maxIter=500))})
  
  # limit lambda to 0 - 1
  if(lambda$modelStruct[[1]] < 0){
    lambda <- gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(0, phy=comp_df[[2]], form = ~Short, fixed=TRUE), 
                 weight = as.formula(N_var), method = "REML",
                 control=list(maxIter=500))
  } else if (lambda$modelStruct[[1]] > 1){
    # this means it is a BM model
     lambda <- gls(as.formula(formulas[[i]]), data = comp_df[[1]], 
                 correlation = corPagel(1, phy=comp_df[[2]], form = ~Short, fixed=TRUE), 
                 weight = as.formula(N_var), method = "REML",
                 control=list(maxIter=500)) 
  }

  corStructure = lambda$modelStruct[[1]]
  mod0 <- gls(as.formula("sperm_length_log_mean ~ 1"), data = comp_df[[1]], 
                 correlation = corPagel(1,
                                        phy=comp_df[[2]], form = ~Short),
                 weight = as.formula(N_var), method = "REML",
                control=list(maxIter=500))
  
  full =   lambda 
  phy_only = mod0
  data_only = lm(as.formula(formulas[[i]]), data = comp_df[[1]])
  R2_pred_data = round(R2.pred(full, phy_only),3)
  R2_pred_phy = round(R2.pred(full, data_only),3)
  R2_pred_full = round(R2.pred(full),3)
  
  #####################
  s_best <- summary(lambda)
  res_df <- data.frame(s_best$tTable)
  res_df$Estimate <- rownames(s_best$tTable)
  res_df$model <- "lambda"
  res <- res_df
  res$R2pred <- c(R2_pred_full, rep(R2_pred_data, nrow(res)-1))
  
  res$predictor = pred_name[[i]]
  res$phylogeny = names(trees)[[j]]
  res$lambda = lambda$modelStruct[[1]]
  res$N = nrow(comp_df[[1]])
  #############
  # DIAGNOSTICS
  if (DIAGNOSTIC == TRUE) {
    png(paste0("fig/diag/",outname, "_", names(trees)[j],"_",preds[i],".png"),
      width = 7, height = 7, unit="in", res = 300)
    pgls_diag_gls(name=paste0(names(trees)[j]," ",predictor), mod_type = "lambda",
                model=lambda, formula=as.formula(formulas[[i]]),
                data = comp_df[[1]], phy=comp_df[[2]], 
                weight = as.formula(N_var), method = "REML", control=list(maxIter=500, opt="optim"))
    dev.off()
  }

  ############
  body_res <- append(body_res,list(res))
  body_res_models <- append(body_res_models, list(lambda))
}
}

body_res <- do.call(rbind, body_res)
body_res$p.value <- format.pval(body_res$p.value, eps=0.001, digits=2)
body_res_out <- body_res %>%
  select(predictor, phylogeny, Estimate, N, lambda, Value, Std.Error, t.value, p.value, R2pred) %>%
  mutate_if(is.numeric, round, 2)
```

```{r}
write.csv(body_res_out, file=paste0("output/PGLS_",outname,"_weighted_PGLS.csv"))

names(body_res_models) <- apply(expand.grid(names(trees), preds), 1, paste, collapse="_")
saveRDS(body_res_models, paste0("output/PGLS_", outname,"_weighted_PGLS.RDS"))
```






